---
title: "Ice_On_Probability_Analysis"
output: html_document
date: "2025-04-14"
---
```{r}
# load libraries
library(pacman)
p_load(dplyr,dataRetrieval,lubridate,tidyr,ggplot2,viridis,readxl,imputeTS,tsibble,sjPlot,pROC,gridExtra,broom,gtsummary,patchwork, stringr)
# This is the file that combines and cleans data:
source("Input_Files/01_Data_Input.R")
```

# Plotting Probability
Now we'll plot the probability of ice as a function of time.
Here we make a new data frame with the imputed values, and add the probability of ice presence given by the model

```{r echo=FALSE}

trimmed_imputed_all_winter <- glm(ice_presence~Flow+temperature_C_impute+cond_uScm_impute+waterYear, data = imputed_data_trimmed_14_23_winter, family = binomial)

probabilities_winter <- predict(trimmed_imputed_all_winter,
  newdata = select(imputed_data_trimmed_14_23_winter, -ice_presence), # remove real outcomes
  type = "response")
imputed_data_trimmed_prob_winter <- imputed_data_trimmed_14_23_winter
imputed_data_trimmed_prob_winter$ice_probability <- probabilities_winter

```

## 0.5 as Threshold
```{r}
# create a vector of predicted probabilities
preds_winter <- predict(trimmed_imputed_all_winter,
  newdata = select(imputed_data_trimmed_14_23_winter, -ice_presence), # remove real outcomes
  type = "response"
)

# if probability < threshold, ice is not on The Loch
preds_outcome_winter <- ifelse(preds_winter < 0.5,
  0,
  1
)

# transform predictions into factor and set labels
preds_outcome_winter <- factor(preds_outcome_winter,
  levels = c(0, 1),
  labels = c("ice", "no ice")
)

# compare observed vs. predicted outcome
tab_winter <- table(imputed_data_trimmed_14_23_winter$ice_presence, preds_outcome_winter,
  dnn = c("observed", "predicted")
)

# print results
tab_winter

accuracy_winter <- sum(diag(tab_winter)) / sum(tab_winter)
accuracy_winter

# sensitivity
sensitivity_winter <- tab_winter[2, 2] / (tab_winter[2, 2] + tab_winter[2, 1])
sensitivity_winter

# specificity
specificity_winter<- tab_winter[1, 1] / (tab_winter[1, 1] + tab_winter[1, 2])
specificity_winter
```


# Ice-On Dates:
Pulling out the first day each year when ice is "ice"
First we have to factor the predicted probability of ice to "ice" and "no ice"
```{r}
winter_probs <- ifelse(probabilities_winter < 0.5,
  1,
  0
)
# changing the labels of the probabilities
winter_probs_factor <- factor(winter_probs, levels = c(0, 1),labels = c("no ice", "ice"))


imputed_data_trimmed_prob_winter$predicted_ice <- winter_probs_factor

```
# This is where I am, and it's broken
```{r}
imputed_data_trimmed_prob_winter <- imputed_data_trimmed_prob_winter %>% filter(waterYear!=2023)
# Create an empty data frame to store the results
ice_on_dates <- data.frame()

# Iterate through each unique waterYear
for(year in unique(imputed_data_trimmed_prob_winter$waterYear)) {
  
  # Filter the data for the current year
  year_data <- imputed_data_trimmed_prob_winter %>% filter(waterYear == year)
  
  # Find the first day when ice_presence == "no ice"
  first_ice_presence <- year_data %>% 
    filter(ice_presence == "ice") %>%
    arrange(wy_doy) %>%
    slice(1)
  
  # Find the first day when predicted_ice == "no ice"
  first_ice_predicted <- year_data %>% 
    filter(predicted_ice == "ice") %>%
    arrange(wy_doy) %>%
    slice(1)
  
  # Create a new row with the results
  result_row <- data.frame(
    waterYear = year,
    first_ice_presence_doy = first_ice_presence$wy_doy,
    first_ice_predicted_doy = first_ice_predicted$wy_doy
  )
  
  # Append the row to the result data frame
  ice_on_dates <- bind_rows(ice_on_dates, result_row)
}

# find the difference between predicted and observed and print the mean of this difference across all years
ice_on_dates$difference <- ((ice_on_dates$first_ice_presence_doy)-(ice_on_dates$first_ice_predicted_doy))*-1
print(mean(ice_on_dates$difference))
# negative means the predicted ice-ff date was EARLIER than the observed


# View the result data frame
#View(imputed_ice_off_dates_predicted_obs)
```

```{r}
ggplot(
  ice_on_dates,
  aes(x = first_ice_presence_doy, y = first_ice_predicted_doy)) +
  theme_bw() +
  geom_point()+
  geom_abline(slope=1,intercept=0)+
  lims(x=c(18,55),y=c(18,55))
```


