---
title: "Ice_On_Probability_Analysis"
output: html_document
date: "2025-04-14"
---
# Load Packages
```{r}
# load libraries
library(pacman)
p_load(dplyr,dataRetrieval,lubridate,tidyr,ggplot2,viridis,readxl,imputeTS,tsibble,sjPlot,pROC,gridExtra,broom,gtsummary,patchwork, stringr, zoo, slider)
# This is the file that combines and cleans data:
source("Input_Files/01_Data_Input.R")
```
# Mutating
## Water Temp Variables
```{r}
# Cumulative water temperature variable:
met_hydro_winter <- met_hydro_winter %>% 
  group_by(waterYear) %>%
  mutate(cumulative_water_temp = cumsum(Temperature_C), 
         wy_doy = seq(1:n())) %>%
  ungroup() %>%
  as.data.frame()

# A rolling slope calculation for the cumulative water temperature variable:
met_hydro_winter <- met_hydro_winter %>%
  group_by(waterYear) %>% 
  arrange(Date) %>%
  mutate(cumulWaterTemp_3day_slope =
           rollapply(cumulative_water_temp,
                     width = 3,
                     FUN = function(x) coef(lm(x ~ seq_along(x)))[2],
                     fill = NA,
                     align = "right")) %>% 
  ungroup() %>% 
  as.data.frame()

met_hydro_winter <- met_hydro_winter %>%
  group_by(waterYear) %>% 
  arrange(Date) %>%
  mutate(cumulWaterTemp_5day_slope =
           rollapply(cumulative_water_temp,
                     width = 5,
                     FUN = function(x) coef(lm(x ~ seq_along(x)))[2],
                     fill = NA,
                     align = "right")) %>% 
  ungroup() %>% 
  as.data.frame()
```
## Air Temp Variables
```{r}
# Create a cumulative freezing degree day variable (CFDD) - cumulative days below 0° using 2m mean air temp
met_hydro_winter <- met_hydro_winter %>%
  group_by(waterYear) %>%
  mutate(
    CFDD = ifelse(
      airT_mean < 0,
      ave(airT_mean < 0,
          cumsum(!(airT_mean < 0) | is.na(airT_mean)),
          FUN = cumsum),
      0
    )
  ) %>%
  ungroup()

# Create a 5-day rolling mean air temperature variable:
met_hydro_winter <- met_hydro_winter %>%
  arrange(Date) %>%  # make sure it's sorted
  mutate(temp_5day_mean = rollmean(airT_mean,
                                    k = 5,
                                    fill = NA,
                                    align = "right"))
# Create a 7-day rolling mean air temperature variable:
met_hydro_winter <- met_hydro_winter %>%
  arrange(Date) %>%  # make sure it's sorted
  mutate(temp_7day_mean = rollmean(airT_mean,
                                    k = 7,
                                    fill = NA,
                                    align = "right"))
ggplot(met_hydro_winter, aes(x = wy_doy, y = temp_7day_mean, color=ice_presence)) +
  geom_point() +
  facet_wrap(~ waterYear, scales = "free_x") +
  theme_minimal()
```
## Conductivity
```{r}
# Adding z score of conductivity to the mix
  met_hydro_winter <- met_hydro_winter %>% 
    group_by(waterYear) %>% 
    mutate(z_cond_uScm=scale(cond_uScm)) %>% 
    ungroup() %>% 
    as.data.frame()
ggplot(met_hydro_winter, aes(x = wy_doy, y = z_cond_uScm, color=ice_presence)) +
  geom_point() +
  facet_wrap(~ waterYear, scales = "free_x") +
  theme_minimal()
```

## Wind Variables
```{r}
# Create a cumulative days with average 10m wind speed below 10mph CLWD (cumulative low wind days)
met_hydro_winter <- met_hydro_winter %>%
  group_by(waterYear) %>%
  mutate(
    CLWD = ifelse(
      wind_10m_mean < 10,
      ave(wind_10m_mean < 10,
          cumsum(!(wind_10m_mean < 10) | is.na(wind_10m_mean)),
          FUN = cumsum),
      0
    )
  ) %>%
  ungroup()

# Create a 2-day rolling mean wind speed variable:
met_hydro_winter <- met_hydro_winter %>%
  arrange(Date) %>%  # make sure it's sorted
  mutate(wind_2day_mean = rollmean(wind_10m_mean,
                                    k = 2,
                                    fill = NA,
                                    align = "right"))

```
# Models
## All and CFDD
```{r echo=FALSE}
# hydro variables stuff
trimmed_imputed_all_winter <- glm(ice_presence~Flow+temperature_C_impute+cond_uScm_impute+waterYear, data = imputed_data_trimmed_14_23_winter, family = binomial)

probabilities_winter <- predict(trimmed_imputed_all_winter,
  newdata = select(imputed_data_trimmed_14_23_winter, -ice_presence), # remove real outcomes
  type = "response")
imputed_data_trimmed_prob_winter <- imputed_data_trimmed_14_23_winter
imputed_data_trimmed_prob_winter$ice_probability <- probabilities_winter

# same thing but with CFDD and CLWD

CFDD_CLWD_model <- glm(ice_presence~CFDD+CLWD+wind_10m_mean+SWin_2m6m_daily_mean+cumulative_dis, data = met_hydro_winter, family = binomial)

CFDD_CLWD_probs <- predict(CFDD_CLWD_model,
  newdata = select(met_hydro_winter, -ice_presence), # remove real outcomes
  type = "response")

met_hydro_winter_probs <- met_hydro_winter
met_hydro_winter_probs$ice_probability <- round(CFDD_CLWD_probs,4)

justprobandice <- met_hydro_winter_probs %>% select(c(ice_probability,ice_presence))
View(justprobandice)

# Table of stats n things:
tab_model(CFDD_CLWD_model,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)

```
## Rolling variables
```{r}
# Rolling mean air temp and cumulative discharge model
rollingAirTemp_cumulDis_model <- glm(ice_presence~temp_7day_mean+cumulative_dis, data = met_hydro_winter, family = binomial)
# Table of stats n things:
tab_model(rollingAirTemp_cumulDis_model,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
# Add those probabilities to the data frame
rollAir_cumDis_probs <- predict(rollingAirTemp_cumulDis_model,
  newdata = select(met_hydro_winter, -ice_presence), # remove real outcomes
  type = "response")

met_hydro_winter_probs$ice_probability_rollAir_cumDis <- round(rollAir_cumDis_probs,4)

# Plot the model probabilities
ggplot(met_hydro_winter_probs, aes(x = wy_doy, y = ice_probability_rollAir_cumDis)) +
  geom_point() +
  facet_wrap(~ waterYear, scales = "free_x") +
  geom_vline(data = ice_on_dates,
             aes(xintercept = first_ice_presence_doy),
             colour = "red",
             linetype = "dashed")+
  theme_minimal()


# Rolling mean air temp model
rollingAirTemp_model <- glm(ice_presence~temp_7day_mean, data = met_hydro_winter, family = binomial)
# Table of stats n things:
tab_model(rollingAirTemp_model,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
# Add those probabilities to the data frame
rollAir_probs <- predict(rollingAirTemp_model,
  newdata = select(met_hydro_winter, -ice_presence), # remove real outcomes
  type = "response")

met_hydro_winter_probs$ice_probability_rollAir <- round(rollAir_probs,4)

# Plot the model probabilities
ggplot(met_hydro_winter_probs, aes(x = wy_doy, y = ice_probability_rollAir, color=ice_presence)) +
  geom_point() +
  facet_wrap(~ waterYear, scales = "free_x") +
  # geom_vline(data = ice_on_dates,
  #            aes(xintercept = first_ice_presence_doy),
  #            colour = "red",
  #            linetype = "dashed")+
  geom_point(aes(y=))
  theme_minimal()
  
  # Rolling mean air temp and z score cond model
rollingAirTempCond_model <- glm(ice_presence~temp_7day_mean+z_cond_uScm, data = met_hydro_winter, family = binomial)
# Table of stats n things:
tab_model(rollingAirTempCond_model,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
# Add those probabilities to the data frame
rollAirCond_probs <- predict(rollingAirTempCond_model,
  newdata = select(met_hydro_winter, -ice_presence), # remove real outcomes
  type = "response")

met_hydro_winter_probs$ice_probability_rollAirCond <- round(rollAirCond_probs,4)

# Plot the model probabilities
ggplot(met_hydro_winter_probs, aes(x = wy_doy, y = ice_probability_rollAirCond, color=ice_presence)) +
  geom_point() +
  facet_wrap(~ waterYear, scales = "free_x") +
  # geom_vline(data = ice_on_dates,
  #            aes(xintercept = first_ice_presence_doy),
  #            colour = "red",
  #            linetype = "dashed")+
  geom_point(aes(y=))
  theme_minimal()
  
```


```{r}
# create a vector of predicted probabilities
preds_ice_on <- predict(rollingAirTemp_model,
  newdata = select(met_hydro_winter, -ice_presence), # remove real outcomes
  type = "response"
)

# if probability < threshold, ice IS on The Loch
preds_outcome_ice_on <- ifelse(preds_ice_on < 0.5,
  1,
  0
)

# transform predictions into factor and set labels
preds_outcome_ice_on <- factor(preds_outcome_ice_on,
  levels = c(1, 0),
  labels = c("ice", "no ice")
)

# compare observed vs. predicted outcome
tab_ice_on <- table(met_hydro_winter$ice_presence, preds_outcome_ice_on,
  dnn = c("observed", "predicted")
)

# print results
tab_ice_on

# accuracy_05 <- sum(diag(tab_05)) / sum(tab_05)
# accuracy_05
# 
# # sensitivity
# sensitivity_05 <- tab_05[2, 2] / (tab_05[2, 2] + tab_05[2, 1])
# sensitivity_05
# 
# # specificity
# specificity_05<- tab_05[1, 1] / (tab_05[1, 1] + tab_05[1, 2])
# specificity_05
```
#Plots
## Daily Air Temp
```{r}
plot_meanAirTemp_year <- function(year) {
  
  ggplot(
    met_hydro_winter %>% 
      filter(waterYear == year, wy_doy > 0),
    aes(x = wy_doy, y = airT_mean, color = factor(waterYear))
  ) +
    theme_bw() +
    geom_point() +
    geom_vline(
      xintercept = ice_on_dates$first_ice_presence_doy[
        ice_on_dates$waterYear == year
      ]
    ) +
    labs(title = paste("Water Year", year),
         x = "Water Year DOY",
         y = "Mean 2m Air Temp",
         color = "Water Year")
}

plot_meanAirTemp_year(2017)
```
## Rolling Air Temp
```{r}
plot_rollingAirTemp_year <- function(year) {
  
  ggplot(
    met_hydro_winter %>% 
      filter(waterYear == year, wy_doy > 0),
    aes(x = wy_doy, y = temp_5day_mean, color = factor(waterYear))
  ) +
    theme_bw() +
    geom_point() +
    geom_vline(
      xintercept = ice_on_dates$first_ice_presence_doy[
        ice_on_dates$waterYear == year
      ]
    ) +
    labs(title = paste("Water Year", year),
         x = "Water Year DOY",
         y = "Rolling 5 Day Mean 2m Air Temp",
         color = "Water Year")
}

plot_rollingAirTemp_year(2015)
```

```{r}
plot_rolling7AirTemp_year <- function(year) {
  
  ggplot(
    met_hydro_winter %>% 
      filter(waterYear == year, wy_doy > 0),
    aes(x = wy_doy, y = temp_7day_mean)
  ) +
    theme_bw() +
    geom_point() +
    geom_vline(
      xintercept = ice_on_dates$first_ice_presence_doy[
        ice_on_dates$waterYear == year
      ]
    )+
    labs(y = "Rolling 7 Day Mean 2m Air Temp",
         color = "Water Year")+
    theme(axis.title.x = element_blank())
}


plot_rolling7AirTemp_year(2014)
```
## Z Conductivity
```{r}
plot_z_cond_year <- function(year) {
  
  ggplot(
    met_hydro_winter %>% 
      filter(waterYear == year, wy_doy > 0),
    aes(x = wy_doy, y = z_cond_uScm)
  ) +
    theme_bw() +
    geom_point() +
    geom_vline(
      xintercept = ice_on_dates$first_ice_presence_doy[
        ice_on_dates$waterYear == year
      ]
    )+
    labs(y = "Z Score Conductivity",
         x = "Water Year DOY",
         color = "Water Year")+
    geom_hline(yintercept = 0,
               linetype = "dashed")
}


plot_z_cond_year(2023)
```

## Rolling Wind Sp
```{r}
plot_rollingWndSpd_year <- function(year) {
  
  ggplot(
    met_hydro_winter %>% 
      filter(waterYear == year, wy_doy > 0),
    aes(x = wy_doy, y = wind_7day_mean, color = factor(waterYear))
  ) +
    theme_bw() +
    geom_point() +
    geom_vline(
      xintercept = ice_on_dates$first_ice_presence_doy[
        ice_on_dates$waterYear == year
      ]
    ) +
    labs(title = paste("Water Year", year),
         x = "Water Year DOY",
         y = "Rolling 7 Day Mean 10m Wind Speed",
         color = "Water Year")
}

plot_rollingWndSpd_year(2016)
```

## Daily Wind Speed
```{r}

plot_wndSpd_year <- function(year) {
  
  ggplot(
    met_hydro_winter %>% 
      filter(waterYear == year, wy_doy > 0),
    aes(x = wy_doy, y = wind_10m_mean, color = factor(waterYear))
  ) +
    theme_bw() +
    geom_point() +
    geom_vline(
      xintercept = ice_on_dates$first_ice_presence_doy[
        ice_on_dates$waterYear == year
      ]
    ) +
    labs(title = paste("Water Year", year),
         x = "Water Year DOY",
         y = "Mean 10m Wind Speed",
         color = "Water Year")
}

plot_wndSpd_year(2017)
```
```{r}
plot_CFDD_year <- function(year) {
  
  ggplot(
    met_hydro_winter %>% 
      filter(waterYear == year, wy_doy > 0),
    aes(x = wy_doy, y = CFDD, color = factor(waterYear))
  ) +
    theme_bw() +
    geom_point() +
    geom_vline(
      xintercept = ice_on_dates$first_ice_presence_doy[
        ice_on_dates$waterYear == year
      ]
    ) +
    labs(title = paste("Water Year", year),
         x = "Water Year DOY",
         y = "Cumulative Freezing Degree Days",
         color = "Water Year")
}

plot_CFDD_year(2015)
```
## Water Temp
```{r}
plot_waterTemp_year <- function(year) {
  
  ggplot(
    met_hydro_winter %>% 
      filter(waterYear == year, wy_doy > 0),
    aes(x = wy_doy, y = Temperature_C, color = factor(waterYear))
  ) +
    theme_bw() +
    geom_point() +
    geom_vline(
      xintercept = ice_on_dates$first_ice_presence_doy[
        ice_on_dates$waterYear == year
      ]
    ) +
    labs(title = paste("Water Year", year),
         x = "Water Year DOY",
         y = "Mean Daily Water Temperature",
         color = "Water Year")
}

plot_waterTemp_year(2016)

# All years of cumulative water temp
ggplot(met_hydro_winter, aes(x = wy_doy, y = cumulative_water_temp)) +
  geom_point() +
  facet_wrap(~ waterYear, scales = "free_x") +
  geom_vline(data = ice_on_dates,
             aes(xintercept = first_ice_presence_doy),
             colour = "red",
             linetype = "dashed")+
  theme_minimal()

# All years of a 3 day slope of cumulative water temp
ggplot(met_hydro_winter, aes(x = wy_doy, y = cumulWaterTemp_3day_slope)) +
  geom_point() +
  facet_wrap(~ waterYear, scales = "free_x") +
  geom_vline(data = ice_on_dates,
             aes(xintercept = first_ice_presence_doy),
             colour = "red",
             linetype = "dashed")+
  theme_minimal()
# All years of a 5 day slope of cumulative water temp
ggplot(met_hydro_winter, aes(x = wy_doy, y = cumulWaterTemp_5day_slope)) +
  geom_point() +
  facet_wrap(~ waterYear, scales = "free_x") +
  geom_vline(data = ice_on_dates,
             aes(xintercept = first_ice_presence_doy),
             colour = "red",
             linetype = "dashed")+
  theme_minimal()
# All years of rolling 7 day air temp

```
## Probability ice-on
```{r}
plot_probability_year <- function(year) {
  
  ggplot(
    prob_winter_daily_data %>% 
      filter(waterYear == year, wy_doy > 0),
    aes(x = wy_doy, y = ice_probability)
  ) +
    theme_bw() +
    geom_point() +
    geom_vline(
      xintercept = ice_on_dates$first_ice_presence_doy[
        ice_on_dates$waterYear == year
      ]
    )+
    labs(title = paste("Water Year", year),
         y = "Probability of No Ice Cover")+
    theme(axis.title.x = element_blank())+
    geom_hline(yintercept = 0.5,
               linetype = "dashed")
}


prob2023 <- plot_probability_year(2023)
```



# Stacked Probability, Rolling Temp, Z Conductivity
```{r}
probTempCondPlot <- prob2023 / rollAir2023 / zCond2023 + plot_layout(heights = c(20,20,20))
probTempCondPlot
ggsave("Figures/Ice_On/probTempCondPlot2023.png", dpi=600, width=6, height=8, units="in")
```

# Adding Met data
# Plotting Probability
Now we'll plot the probability of ice as a function of time.
Here we make a new data frame with the imputed values, and add the probability of ice presence given by the model

```{r echo=FALSE}

probabilities_winter_met <- predict(rollingAirTempCond_model,
  newdata = select(met_hydro_winter, -ice_presence), # remove real outcomes
  type = "response")
prob_winter_daily_data <- met_hydro_winter
prob_winter_daily_data$ice_probability <- probabilities_winter_met

```

## 0.5 as Threshold
```{r}

# if probability < threshold, ice is not on The Loch
preds_outcome_winter_met <- ifelse(probabilities_winter_met < 0.5,
  0,
  1
)

# transform predictions into factor and set labels
preds_outcome_winter_met <- factor(preds_outcome_winter_met,
  levels = c(0, 1),
  labels = c("ice", "no ice")
)

# compare observed vs. predicted outcome
tab_winter_met <- table(prob_winter_daily_data$ice_presence, preds_outcome_winter_met,
  dnn = c("observed", "predicted")
)

# print results
tab_winter_met

accuracy_winter_met <- sum(diag(tab_winter_met)) / sum(tab_winter_met)
accuracy_winter_met

# sensitivity
sensitivity_winter_met <- tab_winter_met[2, 2] / (tab_winter_met[2, 2] + tab_winter_met[2, 1])
sensitivity_winter_met

# specificity
specificity_winter_met<- tab_winter_met[1, 1] / (tab_winter_met[1, 1] + tab_winter_met[1, 2])
specificity_winter_met
```


# Ice-On Dates:
Pulling out the first day each year when ice is "ice"
First we have to factor the predicted probability of ice to "ice" and "no ice"
```{r}
winter_probs_met <- ifelse(probabilities_winter_met < 0.5,
  1,
  0
)
# changing the labels of the probabilities
winter_probs_factor_met <- factor(winter_probs_met, levels = c(0, 1),labels = c("no ice", "ice"))


prob_winter_daily_data$predicted_ice <- winter_probs_factor_met

```
# This is where I am, and it's broken
```{r}
prob_winter_daily_data
# Create an empty data frame to store the results
ice_on_dates <- data.frame()

# Iterate through each unique waterYear
for(year in unique(prob_winter_daily_data$waterYear)) {
  
  # Filter the data for the current year
  year_data <- prob_winter_daily_data %>% filter(waterYear == year)
  
  # Find the first day when ice_presence == "no ice"
  first_ice_presence <- year_data %>% 
    filter(ice_presence == "ice") %>%
    arrange(wy_doy) %>%
    slice(1)
  
  # Find the first day when predicted_ice == "no ice"
  first_ice_predicted <- year_data %>% 
    filter(predicted_ice == "ice") %>%
    arrange(wy_doy) %>%
    slice(1)
  
  # Create a new row with the results
  result_row <- data.frame(
    waterYear = year,
    first_ice_presence_doy = first_ice_presence$wy_doy,
    first_ice_predicted_doy = first_ice_predicted$wy_doy
  )
  
  # Append the row to the result data frame
  ice_on_dates <- bind_rows(ice_on_dates, result_row)
}

# find the difference between predicted and observed and print the mean of this difference across all years
ice_on_dates$difference <- ((ice_on_dates$first_ice_presence_doy)-(ice_on_dates$first_ice_predicted_doy))*-1
print(mean(ice_on_dates$difference))
# negative means the predicted ice-ff date was EARLIER than the observed


# View the result data frame
#View(imputed_ice_off_dates_predicted_obs)
```

```{r}
ggplot(
  ice_on_dates,
  aes(x = first_ice_presence_doy, y = first_ice_predicted_doy)) +
  theme_bw() +
  geom_point()+
  geom_abline(slope=1,intercept=0)+
  lims(x=c(18,55),y=c(18,55))
```