---
title: "Ice_On_Probability_Analysis"
output: html_document
date: "2025-04-14"
---
```{r}
# load libraries
library(pacman)
p_load(dplyr,dataRetrieval,lubridate,tidyr,ggplot2,viridis,readxl,imputeTS,tsibble,sjPlot,pROC,gridExtra,broom,gtsummary,patchwork, stringr)
# This is the file that combines and cleans data:
source("Input_Files/01_Data_Input.R")
```

# Plotting Probability
Now we'll plot the probability of ice as a function of time.
Here we make a new data frame with the imputed values, and add the probability of ice presence given by the model

```{r echo=FALSE}

trimmed_imputed_all_winter <- glm(ice_presence~Flow+temperature_C_impute+cond_uScm_impute+waterYear, data = imputed_data_trimmed_14_23_winter, family = binomial)

probabilities_winter <- predict(trimmed_imputed_all_winter,
  newdata = select(imputed_data_trimmed_14_23_winter, -ice_presence), # remove real outcomes
  type = "response")
imputed_data_trimmed_prob_winter <- imputed_data_trimmed_14_23_winter
imputed_data_trimmed_prob_winter$ice_probability <- probabilities_winter

```

## 0.5 as Threshold
```{r}
# create a vector of predicted probabilities
preds_winter <- predict(trimmed_imputed_all_winter,
  newdata = select(imputed_data_trimmed_14_23_winter, -ice_presence), # remove real outcomes
  type = "response"
)

# if probability < threshold, ice is not on The Loch
preds_outcome_winter <- ifelse(preds_winter < 0.5,
  0,
  1
)

# transform predictions into factor and set labels
preds_outcome_winter <- factor(preds_outcome_winter,
  levels = c(0, 1),
  labels = c("ice", "no ice")
)

# compare observed vs. predicted outcome
tab_winter <- table(imputed_data_trimmed_14_23_winter$ice_presence, preds_outcome_winter,
  dnn = c("observed", "predicted")
)

# print results
tab_winter

accuracy_winter <- sum(diag(tab_winter)) / sum(tab_winter)
accuracy_winter

# sensitivity
sensitivity_winter <- tab_winter[2, 2] / (tab_winter[2, 2] + tab_winter[2, 1])
sensitivity_winter

# specificity
specificity_winter<- tab_winter[1, 1] / (tab_winter[1, 1] + tab_winter[1, 2])
specificity_winter
```


# Ice-On Dates:
Pulling out the first day each year when ice is "ice"
First we have to factor the predicted probability of ice to "ice" and "no ice"
```{r}
winter_probs <- ifelse(probabilities_winter < 0.5,
  1,
  0
)
# changing the labels of the probabilities
winter_probs_factor <- factor(winter_probs, levels = c(0, 1),labels = c("no ice", "ice"))


imputed_data_trimmed_prob_winter$predicted_ice <- winter_probs_factor

```
# This is where I am, and it's broken
```{r}
imputed_data_trimmed_prob_winter <- imputed_data_trimmed_prob_winter %>% filter(waterYear!=2023)
# Create an empty data frame to store the results
ice_on_dates <- data.frame()

# Iterate through each unique waterYear
for(year in unique(imputed_data_trimmed_prob_winter$waterYear)) {
  
  # Filter the data for the current year
  year_data <- imputed_data_trimmed_prob_winter %>% filter(waterYear == year)
  
  # Find the first day when ice_presence == "no ice"
  first_ice_presence <- year_data %>% 
    filter(ice_presence == "ice") %>%
    arrange(wy_doy) %>%
    slice(1)
  
  # Find the first day when predicted_ice == "no ice"
  first_ice_predicted <- year_data %>% 
    filter(predicted_ice == "ice") %>%
    arrange(wy_doy) %>%
    slice(1)
  
  # Create a new row with the results
  result_row <- data.frame(
    waterYear = year,
    first_ice_presence_doy = first_ice_presence$wy_doy,
    first_ice_predicted_doy = first_ice_predicted$wy_doy
  )
  
  # Append the row to the result data frame
  ice_on_dates <- bind_rows(ice_on_dates, result_row)
}

# find the difference between predicted and observed and print the mean of this difference across all years
ice_on_dates$difference <- ((ice_on_dates$first_ice_presence_doy)-(ice_on_dates$first_ice_predicted_doy))*-1
print(mean(ice_on_dates$difference))
# negative means the predicted ice-ff date was EARLIER than the observed


# View the result data frame
#View(imputed_ice_off_dates_predicted_obs)
```

```{r}
ggplot(
  ice_on_dates,
  aes(x = first_ice_presence_doy, y = first_ice_predicted_doy)) +
  theme_bw() +
  geom_point()+
  geom_abline(slope=1,intercept=0)+
  lims(x=c(18,55),y=c(18,55))
```

# Adding Met data
# Plotting Probability
Now we'll plot the probability of ice as a function of time.
Here we make a new data frame with the imputed values, and add the probability of ice presence given by the model

```{r echo=FALSE}

probabilities_winter_met <- predict(daily_winter_met_2m_min_cond_flow,
  newdata = select(met_and_hydro_winter, -ice_presence), # remove real outcomes
  type = "response")
prob_winter_daily_data <- met_and_hydro_winter
prob_winter_daily_data$ice_probability <- probabilities_winter_met

```

## 0.5 as Threshold
```{r}

# if probability < threshold, ice is not on The Loch
preds_outcome_winter_met <- ifelse(probabilities_winter_met < 0.5,
  0,
  1
)

# transform predictions into factor and set labels
preds_outcome_winter_met <- factor(preds_outcome_winter_met,
  levels = c(0, 1),
  labels = c("ice", "no ice")
)

# compare observed vs. predicted outcome
tab_winter_met <- table(prob_winter_daily_data$ice_presence, preds_outcome_winter_met,
  dnn = c("observed", "predicted")
)

# print results
tab_winter_met

accuracy_winter_met <- sum(diag(tab_winter_met)) / sum(tab_winter_met)
accuracy_winter_met

# sensitivity
sensitivity_winter_met <- tab_winter_met[2, 2] / (tab_winter_met[2, 2] + tab_winter_met[2, 1])
sensitivity_winter_met

# specificity
specificity_winter_met<- tab_winter_met[1, 1] / (tab_winter_met[1, 1] + tab_winter_met[1, 2])
specificity_winter_met
```


# Ice-On Dates:
Pulling out the first day each year when ice is "ice"
First we have to factor the predicted probability of ice to "ice" and "no ice"
```{r}
winter_probs_met <- ifelse(probabilities_winter_met < 0.5,
  1,
  0
)
# changing the labels of the probabilities
winter_probs_factor_met <- factor(winter_probs_met, levels = c(0, 1),labels = c("no ice", "ice"))


prob_winter_daily_data$predicted_ice <- winter_probs_factor_met

```
# This is where I am, and it's broken
```{r}
prob_winter_daily_data <- prob_winter_daily_data %>% filter(waterYear!=2023)
# Create an empty data frame to store the results
ice_on_dates_met <- data.frame()

# Iterate through each unique waterYear
for(year in unique(prob_winter_daily_data$waterYear)) {
  
  # Filter the data for the current year
  year_data <- prob_winter_daily_data %>% filter(waterYear == year)
  
  # Find the first day when ice_presence == "no ice"
  first_ice_presence <- year_data %>% 
    filter(ice_presence == "ice") %>%
    arrange(wy_doy) %>%
    slice(1)
  
  # Find the first day when predicted_ice == "no ice"
  first_ice_predicted <- year_data %>% 
    filter(predicted_ice == "ice") %>%
    arrange(wy_doy) %>%
    slice(1)
  
  # Create a new row with the results
  result_row <- data.frame(
    waterYear = year,
    first_ice_presence_doy = first_ice_presence$wy_doy,
    first_ice_predicted_doy = first_ice_predicted$wy_doy
  )
  
  # Append the row to the result data frame
  ice_on_dates_met <- bind_rows(ice_on_dates, result_row)
}

# find the difference between predicted and observed and print the mean of this difference across all years
ice_on_dates_met$difference <- ((ice_on_dates_met$first_ice_presence_doy)-(ice_on_dates_met$first_ice_predicted_doy))*-1
print(mean(ice_on_dates_met$difference))
# negative means the predicted ice-ff date was EARLIER than the observed


# View the result data frame
#View(imputed_ice_off_dates_predicted_obs)
```

```{r}
ggplot(
  ice_on_dates_met,
  aes(x = first_ice_presence_doy, y = first_ice_predicted_doy)) +
  theme_bw() +
  geom_point()+
  geom_abline(slope=1,intercept=0)+
  lims(x=c(18,55),y=c(18,55))
```

# Getting probabilities for the ONLY met model:
```{r}
##### These probabilities are backwards: 1 still represents 100% probability that there's NO ice.. Not sure how to fix this yet

max_met_only_probs <- predict(daily_winter_met_2m_max,
  newdata = select(met_and_hydro_winter, -ice_presence), # remove real outcomes
  type = "response")
max_met_probs_data <- met_and_hydro_winter
max_met_probs_data$ice_probability <- max_met_only_probs


min_met_only_probs <- predict(daily_winter_met_2m_min,
  newdata = select(met_and_hydro_winter, -ice_presence), # remove real outcomes
  type = "response")
min_met_probs_data <- met_and_hydro_winter
min_met_probs_data$ice_probability <- min_met_only_probs

mean_met_only_probs <- predict(daily_winter_met_2m_mean,
  newdata = select(met_and_hydro_winter, -ice_presence), # remove real outcomes
  type = "response")
mean_met_probs_data <- met_and_hydro_winter
mean_met_probs_data$ice_probability <- mean_met_only_probs
```
### Max air temp plots
```{r}
# plots for air temperature and wind speed (change the x variable and limits)
labels_x<-c(1,20,40,60,80)


TempPlot_preddate <- ggplot(
  max_met_probs_data %>%
    filter(waterYear == 2017),
  aes(x = wy_doy, y = T_air_2_m_max, group = waterYear)) +
  theme_bw() +
  geom_line(size=1) +
  geom_point(aes(x = wy_doy, y = T_air_2_m_max, group = waterYear),size=1.5)+
  geom_vline(xintercept = ice_on_dates$first_ice_presence_doy[ice_on_dates$waterYear == 2017], color="forestgreen", size=1)+ # observed ice-on date
  lims(x = c(1, 92), y = c(0, 10))+
  theme(axis.title.x = element_blank())+
  labs(y = "2m Max Daily Air Temperature (°C)")+
  scale_x_continuous(breaks=labels_x,labels=c("Oct-01","Oct-20","Nov-09","Nov-29","Dec-19"),limits=c(1,92))

windPlot_preddate <- ggplot(
  max_met_probs_data %>%
    filter(waterYear == 2017),
  aes(x = wy_doy, y = WSpd_2_m_mean, group = waterYear)) +
  theme_bw() +
  geom_line(size=1) +
  geom_point(aes(x = wy_doy, y = WSpd_2_m_mean, group = waterYear),size=1.5)+
  geom_vline(xintercept = ice_on_dates$first_ice_presence_doy[ice_on_dates$waterYear == 2017], color="forestgreen", size=1)+ # observed ice-on date
  lims(x = c(1, 92), y = c(0, 15))+
  theme(axis.title.x = element_blank())+
  labs(y = "2m Average Daily Wind Speed m/s")+
  scale_x_continuous(breaks=labels_x,labels=c("Oct-01","Oct-20","Nov-09","Nov-29","Dec-19"),limits=c(1,92))


probIcePlot <- ggplot(
  max_met_probs_data %>%
    filter(waterYear == 2017),
  aes(x = wy_doy, y = ice_probability, group = waterYear)) +
  theme_bw() +
  geom_line(size=1) +
  geom_point(aes(x = wy_doy, y = ice_probability, group = waterYear),size=1.5)+
  geom_vline(xintercept = ice_on_dates$first_ice_presence_doy[ice_on_dates$waterYear == 2017], color="forestgreen", size=1)+ # observed ice-on date
  lims(x = c(1, 92), y = c(0, 1))+
  theme(axis.title.x = element_blank())+
  labs(y = "Probability of No Ice")+
  scale_x_continuous(breaks=labels_x,labels=c("Oct-01","Oct-20","Nov-09","Nov-29","Dec-19"),limits=c(1,92))


TempPlot_preddate
windPlot_preddate
probIcePlot


probTempWindPlot <- probIcePlot / TempPlot_preddate / windPlot_preddate + plot_layout(heights = c(20,20,20))
probTempWindPlot

ggsave("Figures/MAXprobTempWind2017.png", dpi=600, width=6, height=8, units="in")
```
### Min air temp plots

```{r}
# plots for air temperature and wind speed (change the x variable and limits)
labels_x<-c(1,20,40,60,80)


TempPlot_preddate <- ggplot(
  min_met_probs_data %>%
    filter(waterYear == 2017),
  aes(x = wy_doy, y = T_air_2_m_min, group = waterYear)) +
  theme_bw() +
  geom_line(size=1) +
  geom_point(aes(x = wy_doy, y = T_air_2_m_max, group = waterYear),size=1.5)+
  geom_vline(xintercept = ice_on_dates$first_ice_presence_doy[ice_on_dates$waterYear == 2017], color="forestgreen", size=1)+ # observed ice-on date
  lims(x = c(1, 92), y = c(0, 10))+
  theme(axis.title.x = element_blank())+
  labs(y = "2m Min Daily Air Temperature (°C)")+
  scale_x_continuous(breaks=labels_x,labels=c("Oct-01","Oct-20","Nov-09","Nov-29","Dec-19"),limits=c(1,92))

windPlot_preddate <- ggplot(
  min_met_probs_data %>%
    filter(waterYear == 2017),
  aes(x = wy_doy, y = WSpd_2_m_mean, group = waterYear)) +
  theme_bw() +
  geom_line(size=1) +
  geom_point(aes(x = wy_doy, y = WSpd_2_m_mean, group = waterYear),size=1.5)+
  geom_vline(xintercept = ice_on_dates$first_ice_presence_doy[ice_on_dates$waterYear == 2017], color="forestgreen", size=1)+ # observed ice-on date
  lims(x = c(1, 92), y = c(0, 15))+
  theme(axis.title.x = element_blank())+
  labs(y = "2m Average Daily Wind Speed m/s")+
  scale_x_continuous(breaks=labels_x,labels=c("Oct-01","Oct-20","Nov-09","Nov-29","Dec-19"),limits=c(1,92))


probIcePlot <- ggplot(
  min_met_probs_data %>%
    filter(waterYear == 2017),
  aes(x = wy_doy, y = ice_probability, group = waterYear)) +
  theme_bw() +
  geom_line(size=1) +
  geom_point(aes(x = wy_doy, y = ice_probability, group = waterYear),size=1.5)+
  geom_vline(xintercept = ice_on_dates$first_ice_presence_doy[ice_on_dates$waterYear == 2017], color="forestgreen", size=1)+ # observed ice-on date
  lims(x = c(1, 92), y = c(0, 1))+
  theme(axis.title.x = element_blank())+
  labs(y = "Probability of No Ice")+
  scale_x_continuous(breaks=labels_x,labels=c("Oct-01","Oct-20","Nov-09","Nov-29","Dec-19"),limits=c(1,92))


TempPlot_preddate
windPlot_preddate
probIcePlot


probTempWindPlot <- probIcePlot / TempPlot_preddate / windPlot_preddate + plot_layout(heights = c(20,20,20))
probTempWindPlot

ggsave("Figures/MINprobTempWind2017.png", dpi=600, width=6, height=8, units="in")
```
### Mean air temp plots
```{r}
# plots for air temperature and wind speed (change the x variable and limits)
labels_x<-c(1,20,40,60,80)


TempPlot_preddate <- ggplot(
  mean_met_probs_data %>%
    filter(waterYear == 2014),
  aes(x = wy_doy, y = T_air_2_m_mean, group = waterYear)) +
  theme_bw() +
  geom_line(size=1) +
  geom_point(aes(x = wy_doy, y = T_air_2_m_max, group = waterYear),size=1.5)+
  geom_vline(xintercept = ice_on_dates$first_ice_presence_doy[ice_on_dates$waterYear == 2014], color="forestgreen", size=1)+ # observed ice-on date
  lims(x = c(1, 92), y = c(0, 10))+
  theme(axis.title.x = element_blank())+
  labs(y = "2m Mean Daily Air Temperature (°C)")+
  scale_x_continuous(breaks=labels_x,labels=c("Oct-01","Oct-20","Nov-09","Nov-29","Dec-19"),limits=c(1,92))

windPlot_preddate <- ggplot(
  mean_met_probs_data %>%
    filter(waterYear == 2014),
  aes(x = wy_doy, y = WSpd_2_m_mean, group = waterYear)) +
  theme_bw() +
  geom_line(size=1) +
  geom_point(aes(x = wy_doy, y = WSpd_2_m_mean, group = waterYear),size=1.5)+
  geom_vline(xintercept = ice_on_dates$first_ice_presence_doy[ice_on_dates$waterYear == 2014], color="forestgreen", size=1)+ # observed ice-on date
  lims(x = c(1, 92), y = c(0, 15))+
  theme(axis.title.x = element_blank())+
  labs(y = "2m Average Daily Wind Speed m/s")+
  scale_x_continuous(breaks=labels_x,labels=c("Oct-01","Oct-20","Nov-09","Nov-29","Dec-19"),limits=c(1,92))


probIcePlot <- ggplot(
  mean_met_probs_data %>%
    filter(waterYear == 2014),
  aes(x = wy_doy, y = ice_probability, group = waterYear)) +
  theme_bw() +
  geom_line(size=1) +
  geom_point(aes(x = wy_doy, y = ice_probability, group = waterYear),size=1.5)+
  geom_vline(xintercept = ice_on_dates$first_ice_presence_doy[ice_on_dates$waterYear == 2014], color="forestgreen", size=1)+ # observed ice-on date
  lims(x = c(1, 92), y = c(0, 1))+
  theme(axis.title.x = element_blank())+
  labs(y = "Probability of No Ice")+
  scale_x_continuous(breaks=labels_x,labels=c("Oct-01","Oct-20","Nov-09","Nov-29","Dec-19"),limits=c(1,92))


TempPlot_preddate
windPlot_preddate
probIcePlot


probTempWindPlot <- probIcePlot / TempPlot_preddate / windPlot_preddate + plot_layout(heights = c(20,20,20))
probTempWindPlot

ggsave("Figures/MEANprobTempWind2014.png", dpi=600, width=6, height=8, units="in")
```