---
title: "New Models"
output: html_document
date: "2025-02-20"
---
```{r}
# load libraries
library(pacman)
p_load(dplyr,dataRetrieval,lubridate,tidyr,ggplot2,viridis,readxl,imputeTS,tsibble,sjPlot,pROC,gridExtra,broom,gtsummary)
# This is the file that combines and cleans data:
source("Input_Files/01_Data_Input.R")
```

# Ice-Off Models:
Here we start to make simple binary logistic regression models with different variables and compare their accuracy.

April 1 - July 15 Trimmed vs. Un-Trimmed Data - all variables:
These models take into account all the variables: temperature, conductivity, flow, cumulative discharge, and water year.
```{r}
# untrimmed data:
untrimmed_daily_all <- glm(ice_presence~Flow+cumulative_dis+Temperature_C+cond_uScm+waterYear, data = flow_temp_cond_daily_ice, family = binomial)

untrimmed_weekly_all <- glm(ice_presence~Flow+cumulative_dis+temperature_C_weekly+cond_uScm_weekly+waterYear, data = flow_temp_cond_weekly_ice_14_23, family = binomial)

untrimmed_imputed_all <- glm(ice_presence~Flow+cumulative_dis+temperature_C_impute+cond_uScm_impute+waterYear, data = flow_temp_cond_imputed_ice_14_23, family = binomial)

# trimmed data:
trimmed_daily_all <- glm(ice_presence~Flow+cumulative_dis+Temperature_C+cond_uScm+waterYear, data = daily_data_trimmed, family = binomial)

trimmed_weekly_all <- glm(ice_presence~Flow+cumulative_dis+temperature_C_weekly+cond_uScm_weekly+waterYear, data = weekly_data_trimmed_14_23, family = binomial)

trimmed_imputed_all <- glm(ice_presence~Flow+cumulative_dis+temperature_C_impute+cond_uScm_impute+waterYear, data = imputed_data_trimmed_14_23, family = binomial)

```
How do they compare?
The R^2 for the models using data trimmed to include only April 1 - July 15 increase by ~ 0.2 for each different temporal frequency. So from now on, we will be using the trimmed data.
```{r}
# supplemental table with all models
tab_model(untrimmed_daily_all,untrimmed_weekly_all,untrimmed_imputed_all,trimmed_daily_all,trimmed_imputed_all,trimmed_weekly_all,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
# reduced table for main body of thesis
tab_model(trimmed_daily_all,trimmed_imputed_all,trimmed_weekly_all,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
```

Now that we know to use the trimmed data, we'll narrow down the variables used in the model. We'll start by removing flow because cumulative discharge had a lower p-value in the trimmed models.
```{r}
trimmed_daily_reduced1a <- glm(ice_presence~cumulative_dis+Temperature_C+cond_uScm+waterYear, data = daily_test_df, family = binomial)

trimmed_weekly_reduced1a <- glm(ice_presence~cumulative_dis+temperature_C_weekly+cond_uScm_weekly+waterYear, data = weekly_test_df, family = binomial)

trimmed_imputed_reduced1a <- glm(ice_presence~cumulative_dis+temperature_C_impute+cond_uScm_impute+waterYear, data = imputed_test_df, family = binomial)
```
Results: <0.01 change in R^2
```{r}
tab_model(trimmed_daily_all, trimmed_imputed_all,trimmed_weekly_all, trimmed_daily_reduced1a,trimmed_imputed_reduced1a,trimmed_weekly_reduced1a,
  show.intercept = FALSE,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
```
Now we'll remove water year because it was not significant in any of the reduced models.
```{r}
trimmed_daily_reduced2a <- glm(ice_presence~cumulative_dis+Temperature_C+cond_uScm, data = daily_test_df, family = binomial)

trimmed_weekly_reduced2a <- glm(ice_presence~cumulative_dis+temperature_C_weekly+cond_uScm_weekly, data = weekly_test_df, family = binomial)

trimmed_imputed_reduced2a <- glm(ice_presence~cumulative_dis+temperature_C_impute+cond_uScm_impute, data = imputed_test_df, family = binomial)
```
Results: No change in R^2
```{r}
tab_model(trimmed_daily_reduced2a, trimmed_imputed_reduced2a,trimmed_weekly_reduced2a, trimmed_daily_reduced1a,trimmed_imputed_reduced1a,trimmed_weekly_reduced1a,
  show.intercept = FALSE,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
```
Now we'll remove conductivity because its p-value was only <0.05 in one model. These are the final models:
```{r}
daily_final_model_test <- glm(ice_presence~cumulative_dis+Temperature_C, data = daily_test_df, family = binomial)
# trimmed_daily_reduced_final
weekly_final_model_test <- glm(ice_presence~cumulative_dis+temperature_C_weekly, data = weekly_test_df, family = binomial)
#trimmed_weekly_reduced_final
imputed_final_model_test <- glm(ice_presence~cumulative_dis+temperature_C_impute, data = imputed_test_df, family = binomial)
#trimmed_imputed_reduced_final


# Also run the final model with the validation time period to be able to compare the R^2:
imputed_validation_model <- glm(ice_presence~cumulative_dis+temperature_C_impute, data = imputed_validation_df, family = binomial)
```
Results: -0.011 change for weekly R^2, <0.01 change in daily and imputed R^2
```{r}
tab_model(daily_final_model_test, imputed_final_model_test,weekly_final_model_test, trimmed_daily_reduced2a,trimmed_imputed_reduced2a,trimmed_weekly_reduced2a,
  show.intercept = FALSE,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)

# FOR THESIS DOCUMENT:
# final three different temporal resolution models with the validation model:
tab_model(daily_final_model_test, imputed_final_model_test,weekly_final_model_test,imputed_validation_model,
  show.intercept = FALSE,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
```

# Writing the model formula
```{r}
# Extracting the coefficients

coefficients <- coef(imputed_final_model_test) 

# Construct the equation string

equation <- paste("Ice Presence =", round(coefficients[1], 2), " +", 

                 round(coefficients[2], 2), "* Cumulative Discharge +", 

                 round(coefficients[3], 2), "* Water Temperature") 

print(equation) 
```




Now that we have models that explain ~86% of the data (imputed model), we'll make some plots to determine the thresholds at which the models should determine ice-off - Now go to the Next Rmd ("02_Probability_Ice_Analysis.Rmd")
```{r}
discharge_plot <- plot_model(imputed_final_model_test,
  type = "pred",
  terms = c("cumulative_dis [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")

temp_plot <- plot_model(imputed_final_model_test,
  type = "pred",
  terms = c("temperature_C_impute [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")

discharge_plot
temp_plot
```





# Ice-On Models:
Here we start to make simple binary logistic regression models with different variables and compare their accuracy.

October 1 - December 31 Trimmed
These models take into account all the variables: temperature, conductivity, flow, cumulative discharge, and water year.
```{r}

# trimmed for winter data:
trimmed_daily_all_winter <- glm(ice_presence~Flow+cumulative_dis+Temperature_C+cond_uScm+waterYear, data = daily_data_trimmed_winter, family = binomial)

trimmed_weekly_all_winter <- glm(ice_presence~Flow+cumulative_dis+temperature_C_weekly+cond_uScm_weekly+waterYear, data = weekly_data_trimmed_14_23_winter, family = binomial)

trimmed_imputed_all_winter <- glm(ice_presence~Flow+cumulative_dis+temperature_C_impute+cond_uScm_impute+waterYear, data = imputed_data_trimmed_14_23_winter, family = binomial)

```
How do they compare?
The R^2 for the models using data trimmed to include only April 1 - July 15 increase by ~ 0.2 for each different temporal frequency. So from now on, we will be using the trimmed data.
```{r}
# supplemental table with all models
tab_model(trimmed_daily_all_winter,trimmed_imputed_all_winter,trimmed_weekly_all_winter,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  show.intercept = FALSE,
  p.style = "numeric_stars" # display p-values and stars
)
```

Now that we know to use the trimmed data, we'll narrow down the variables used in the model. We'll start by removing flow because cumulative discharge had a lower p-value in the trimmed models.
```{r}
trimmed_daily_reduced1a_winter <- glm(ice_presence~Flow+Temperature_C+cond_uScm+waterYear, data = daily_data_trimmed_winter, family = binomial)

trimmed_weekly_reduced1a_winter <- glm(ice_presence~Flow+temperature_C_weekly+cond_uScm_weekly+waterYear, data = weekly_data_trimmed_14_23_winter, family = binomial)

trimmed_imputed_reduced1a_winter <- glm(ice_presence~Flow+temperature_C_impute+cond_uScm_impute+waterYear, data = imputed_data_trimmed_14_23_winter, family = binomial)
```

```{r}
tab_model(trimmed_daily_all_winter, trimmed_imputed_all_winter,trimmed_weekly_all_winter, trimmed_daily_reduced1a_winter,trimmed_imputed_reduced1a_winter,trimmed_weekly_reduced1a_winter,
  show.intercept = FALSE,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
```

```{r}
trimmed_daily_reduced2a_winter <- glm(ice_presence~Flow+Temperature_C+waterYear, data = daily_data_trimmed_winter, family = binomial)

trimmed_weekly_reduced2a_winter <- glm(ice_presence~Flow+temperature_C_weekly+waterYear, data = weekly_data_trimmed_14_23_winter, family = binomial)

trimmed_imputed_reduced2a_winter <- glm(ice_presence~Flow+temperature_C_impute+waterYear, data = imputed_data_trimmed_14_23_winter, family = binomial)
```

```{r}
tab_model(trimmed_daily_all_winter, trimmed_imputed_all_winter,trimmed_weekly_all_winter, trimmed_daily_reduced1a_winter,trimmed_imputed_reduced1a_winter,trimmed_weekly_reduced1a_winter,trimmed_daily_reduced2a_winter,trimmed_imputed_reduced2a_winter,trimmed_weekly_reduced2a_winter,
  show.intercept = FALSE,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
```
## Met Models
```{r}
# 2 meter air temp daily min
daily_winter_met_2m_min <- glm(ice_presence~T_air_2_m_min+WSpd_2_m_mean+waterYear, data = met_and_hydro_winter, family = binomial)

tab_model(daily_winter_met_2m_min,
  show.intercept = FALSE,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)

# 2  meter air temp daily max
daily_winter_met_2m_max <- glm(ice_presence~T_air_2_m_max+WSpd_2_m_mean+waterYear, data = met_and_hydro_winter, family = binomial)

tab_model(daily_winter_met_2m_max,
  show.intercept = FALSE,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)

# 2  meter air temp daily mean
daily_winter_met_2m_mean <- glm(ice_presence~T_air_2_m_mean+WSpd_2_m_mean+waterYear, data = met_and_hydro_winter, family = binomial)

tab_model(daily_winter_met_2m_mean,
  show.intercept = FALSE,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)

# 2  meter air temp daily min and conductivity
daily_winter_met_2m_min_cond <- glm(ice_presence~T_air_2_m_min+WSpd_2_m_mean+cond_uScm+Flow+waterYear, data = met_and_hydro_winter, family = binomial)

tab_model(daily_winter_met_2m_min_cond,
  show.intercept = FALSE,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
```

