---
title: "Models"
output: html_document
date: "2024-11-23"
---
```{r}
# load libraries
library(pacman)
p_load(dplyr,dataRetrieval,lubridate,tidyr,ggplot2,viridis,readxl,imputeTS,tsibble,sjPlot,pROC,gridExtra,broom,gtsummary)
# This is the file that combines and cleans data:
source("Input_Files/01_Data_Input.R")
```









# Models:
Here we start to make simple binary logistic regression models with different variables and compare their accuracy.

April 1 - July 15 Trimmed vs. Un-Trimmed Data - all variables:
These models take into account all the variables: temperature, conductivity, flow, cumulative discharge, and water year.
```{r}
# untrimmed data:
untrimmed_daily_all <- glm(ice_presence~Flow+cumulative_dis+Temperature_C+cond_uScm+waterYear, data = flow_temp_cond_daily_ice, family = binomial)

untrimmed_weekly_all <- glm(ice_presence~Flow+cumulative_dis+temperature_C_raw+cond_uScm+waterYear, data = flow_temp_cond_weekly_ice_14_23, family = binomial)

untrimmed_imputed_all <- glm(ice_presence~Flow+cumulative_dis+temperature_C_impute+cond_uScm_impute+waterYear, data = flow_temp_cond_imputed_ice_14_23, family = binomial)

# trimmed data:
trimmed_daily_all <- glm(ice_presence~Flow+cumulative_dis+Temperature_C+cond_uScm+waterYear, data = daily_data_trimmed, family = binomial)

trimmed_weekly_all <- glm(ice_presence~Flow+cumulative_dis+temperature_C_raw+cond_uScm+waterYear, data = weekly_data_trimmed_14_23, family = binomial)

trimmed_imputed_all <- glm(ice_presence~Flow+cumulative_dis+temperature_C_impute+cond_uScm_impute+waterYear, data = imputed_data_trimmed_14_23, family = binomial)

```
How do they compare?
The R^2 for the models using data trimmed to include only April 1 - July 15 increase by ~ 0.2 for each different temporal frequency. So from now on, we will be using the trimmed data.
```{r}
tab_model(untrimmed_daily_all,untrimmed_weekly_all,untrimmed_imputed_all,trimmed_daily_all,trimmed_imputed_all,trimmed_weekly_all,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
```
Now that we know to use the trimmed data, we'll narrow down the variables used in the model. We'll start by removing flow because we have cumulative discharge had a lower p-value in the trimmed models.
```{r}
trimmed_daily_reduced1 <- glm(ice_presence~cumulative_dis+Temperature_C+cond_uScm+waterYear, data = daily_data_trimmed, family = binomial)

trimmed_weekly_reduced1 <- glm(ice_presence~cumulative_dis+temperature_C_raw+cond_uScm+waterYear, data = weekly_data_trimmed_14_23, family = binomial)

trimmed_imputed_reduced1 <- glm(ice_presence~cumulative_dis+temperature_C_impute+cond_uScm_impute+waterYear, data = imputed_data_trimmed_14_23, family = binomial)
```
Results: <0.01 change in R^2
```{r}
tab_model(trimmed_daily_reduced1, trimmed_imputed_reduced1,trimmed_weekly_reduced1, trimmed_daily_all,trimmed_imputed_all,trimmed_weekly_all,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
```
Now we'll remove water year because it was not significant in any of the reduced models.
```{r}
trimmed_daily_reduced2 <- glm(ice_presence~cumulative_dis+Temperature_C+cond_uScm, data = daily_data_trimmed, family = binomial)

trimmed_weekly_reduced2 <- glm(ice_presence~cumulative_dis+temperature_C_raw+cond_uScm, data = weekly_data_trimmed_14_23, family = binomial)

trimmed_imputed_reduced2 <- glm(ice_presence~cumulative_dis+temperature_C_impute+cond_uScm_impute, data = imputed_data_trimmed_14_23, family = binomial)
```
Results: No change in R^2
```{r}
tab_model(trimmed_daily_reduced2, trimmed_imputed_reduced2,trimmed_weekly_reduced2, trimmed_daily_reduced1,trimmed_imputed_reduced1,trimmed_weekly_reduced1,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
```
Now we'll remove conductivity because its p-value was only <0.05 in one model
```{r}
trimmed_daily_reduced_final <- glm(ice_presence~cumulative_dis+Temperature_C, data = daily_data_trimmed, family = binomial)

trimmed_weekly_reduced_final <- glm(ice_presence~cumulative_dis+temperature_C_raw, data = weekly_data_trimmed_14_23, family = binomial)

trimmed_imputed_reduced_final <- glm(ice_presence~cumulative_dis+temperature_C_impute, data = imputed_data_trimmed_14_23, family = binomial)
```
Results: -0.011 change for weekly R^2, <0.01 change in daily and imputed R^2
```{r}
tab_model(trimmed_daily_reduced_final, trimmed_imputed_reduced_final,trimmed_weekly_reduced_final, trimmed_daily_reduced2,trimmed_imputed_reduced2,trimmed_weekly_reduced2,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
```

# Functional Ice-off Dates
```{r}
# here, the model using the first date of observed 0% ice as the ice-off date in the training data is compared to the model using the date of "functional ice-off" in the training data.
tab_model(trimmed_imputed_reduced_final,trimmed_imputed_reduced_final_func,
  show.ci = FALSE, # remove CI
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
```


Now that we have models that explain ~86% of the data (imputed model), we'll make some plots to determine the thresholds at which the models should determine ice-off - Now go to the Next Rmd ("01_