---
title: "Plotting Probability"
output: html_document
date: "2024-11-23"
---

```{r}
# load libraries
library(pacman)
p_load(dplyr,dataRetrieval,lubridate,tidyr,ggplot2,viridis,readxl,imputeTS,tsibble,sjPlot,pROC,gridExtra,broom,gtsummary,patchwork, stringr)
# This is the file that combines and cleans data:
source("Input_Files/01_Data_Input.R")
```
# Plotting Probability
Now we'll plot the probability of ice as a function of time.
Here we make a new data frame with the imputed values, and add the probability of ice presence given by the model

*Keep in mind that a probability of 1 = NO ICE and a probability of 0 = ICE. (Because we are predicting ice-off dates)
```{r echo=FALSE}

probabilities_impute <- predict(trimmed_imputed_reduced_final,
  newdata = select(imputed_data_trimmed_14_23, -ice_presence), # remove real outcomes
  type = "response")
imputed_data_trimmed_prob <- imputed_data_trimmed_14_23
imputed_data_trimmed_prob$ice_probability <- probabilities_impute

```
# Cumul. Discharge and Probability of Ice
## 2014:
```{r echo=FALSE}

plot2014 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2014), aes(x = wy_doy, y = ice_probability)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=obs_ice_melt_windows$wy_doy_100_ice[obs_ice_melt_windows$waterYear == 2014], xmax=obs_ice_melt_windows$wy_doy_0_ice[obs_ice_melt_windows$waterYear == 2014], ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  labs(title = "Predicted Probability of Ice Over Time",x = "Water Year Day of Year",y = "Probability of Ice")+ 
  theme_dark()+
  lims(x = c(200,265))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))

plot2014_2 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2014), aes(x = wy_doy, y = cumulative_dis)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=obs_ice_melt_windows$wy_doy_100_ice[obs_ice_melt_windows$waterYear == 2014], xmax=obs_ice_melt_windows$wy_doy_0_ice[obs_ice_melt_windows$waterYear == 2014], ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  labs(
    title = "Cumulative Discharge Over Time",
    x = "Water Year Day of Year",
    y = "Cumulative Discharge (cfs)")+ lims(x = c(200,265)) + 
    theme_dark()+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))

plot2014_3 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2014), aes(x = wy_doy, y = temperature_C_impute)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=obs_ice_melt_windows$wy_doy_100_ice[obs_ice_melt_windows$waterYear == 2014], xmax=obs_ice_melt_windows$wy_doy_0_ice[obs_ice_melt_windows$waterYear == 2014], ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark()+
  labs(
    title = "Temperature Over Time",
    x = "Water Year Day of Year",
    y = "Temperature (C)"
  ) + lims(x = c(200,265))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))

#Use patchwork library to annotate plots and combine -- I find it easier to use than gridExtra or cowplot
WY2014_all <- (plot2014 / plot2014_2 / plot2014_3) + plot_annotation(
  title = 'WY 2014',
  subtitle = str_wrap('Shaded area indicates windows from weekly photographs (full cover to no coverage)', width=50)
)
WY2014_all
```
## 2015:
```{r echo=FALSE}

plot2015 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2015), aes(x = wy_doy, y = ice_probability)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=obs_ice_melt_windows$wy_doy_100_ice[obs_ice_melt_windows$waterYear == 2015], xmax=obs_ice_melt_windows$wy_doy_0_ice[obs_ice_melt_windows$waterYear == 2015], ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  labs(
    title = "Predicted Probability of Ice Over Time",
    x = "Water Year Day of Year",
    y = "Probability of Ice"
  )+ lims(x = c(195,265))+
  theme_dark()+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))

plot2015_2 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2015), aes(x = wy_doy, y = cumulative_dis)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=obs_ice_melt_windows$wy_doy_100_ice[obs_ice_melt_windows$waterYear == 2015], xmax=obs_ice_melt_windows$wy_doy_0_ice[obs_ice_melt_windows$waterYear == 2015], ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  labs(
    title = "Cumulative Discharge Over Time",
    x = "Water Year Day of Year",
    y = "Cumulative Discharge (cfs)") + 
    lims(x = c(195,265))+
    theme_dark()+
    theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))

grid.arrange(plot2015, plot2015_2, ncol = 1)
```
## 2016:
```{r echo=FALSE}

plot2016 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2016), aes(x = wy_doy, y = ice_probability)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=195,xmax=251,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  labs(
    title = "Predicted Probability of Ice Over Time",
    x = "Water Year Day of Year",
    y = "Probability of Ice"
  )+ lims(x = c(190,270))+
  theme_dark()+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))

plot2016_2 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2016), aes(x = wy_doy, y = cumulative_dis)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=195,xmax=251,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  labs(
    title = "Cumulative Discharge Over Time",
    x = "Water Year Day of Year",
    y = "Cumulative Discharge (cfs)"
  ) + lims(x = c(190,270))+
  theme_dark()+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))

grid.arrange(plot2016, plot2016_2, ncol = 1)
```
## 2017:
```{r fig.height}

plot2017 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2017), aes(x = wy_doy, y = ice_probability)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=173,xmax=243,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  labs(
    title = "Predicted Probability of Ice Over Time",
    x = "Water Year Day of Year",
    y = "Probability of Ice"
  )+ lims(x = c(170,260))+
  theme_dark()+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))

plot2017_2 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2017), aes(x = wy_doy, y = cumulative_dis)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=173,xmax=243,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  labs(
    title = "Cumulative Discharge Over Time",
    x = "Water Year Day of Year",
    y = "Cumulative Discharge (cfs)"
  )+ lims(x = c(170,260)) +
  theme_dark()+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))

grid.arrange(plot2017, plot2017_2, ncol = 1)
```
## 2018:
```{r echo=FALSE}

plot2018 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2018), aes(x = wy_doy, y = ice_probability)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=214,xmax=235,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Predicted Probability of Ice Over Time",
    x = "Water Year Day of Year",
    y = "Probability of Ice"
  )+ lims(x = c(200,250))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))

plot2018_2 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2018), aes(x = wy_doy, y = cumulative_dis)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=214,xmax=235,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Cumulative Discharge Over Time",
    x = "Water Year Day of Year",
    y = "Cumulative Discharge (cfs)"
  ) + lims(x = c(200,250))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))

grid.arrange(plot2018, plot2018_2, ncol = 1)
```
## 2019:
```{r echo=FALSE}

plot2019 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2019), aes(x = wy_doy, y = ice_probability)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=206,xmax=260,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Predicted Probability of Ice Over Time",
    x = "Water Year Day of Year",
    y = "Probability of Ice"
  )+ lims(x = c(200,270))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))
plot2019_2 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2019), aes(x = wy_doy, y = cumulative_dis)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=206,xmax=260,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Cumulative Discharge Over Time",
    x = "Water Year Day of Year",
    y = "Cumulative Discharge (cfs)"
  ) + lims(x = c(200,270))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))
grid.arrange(plot2019, plot2019_2, ncol = 1)
```
## 2020:
```{r echo=FALSE}

plot2020 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2020), aes(x = wy_doy, y = ice_probability)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=211,xmax=234,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Predicted Probability of Ice Over Time",
    x = "Water Year Day of Year",
    y = "Probability of Ice"
  )+ lims(x = c(200,250))+ 
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))
plot2020_2 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2020), aes(x = wy_doy, y = cumulative_dis)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=211,xmax=234,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Cumulative Discharge Over Time",
    x = "Water Year Day of Year",
    y = "Cumulative Discharge (cfs)"
  ) + lims(x = c(200,250))+ 
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))
grid.arrange(plot2020, plot2020_2, ncol = 1)
```
## 2021:
```{r echo=FALSE}

plot2021 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2021), aes(x = wy_doy, y = ice_probability)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=210,xmax=255,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Predicted Probability of Ice Over Time",
    x = "Water Year Day of Year",
    y = "Probability of Ice"
  )+ lims(x = c(200,270))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))

plot2021_2 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2021), aes(x = wy_doy, y = cumulative_dis)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=210,xmax=255,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Cumulative Discharge Over Time",
    x = "Water Year Day of Year",
    y = "Cumulative Discharge (cfs)"
  ) + lims(x = c(200,270))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))
grid.arrange(plot2021, plot2021_2, ncol = 1)
```
## 2022:
```{r echo=FALSE}

plot2022 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2022), aes(x = wy_doy, y = ice_probability)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=216,xmax=244,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Predicted Probability of Ice Over Time",
    x = "Water Year Day of Year",
    y = "Probability of Ice"
  )+ lims(x = c(210,270))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))

plot2022_2 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2022), aes(x = wy_doy, y = cumulative_dis)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=216,xmax=244,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Cumulative Discharge Over Time",
    x = "Water Year Day of Year",
    y = "Cumulative Discharge (cfs)"
  ) + lims(x = c(210,270))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))
grid.arrange(plot2022, plot2022_2, ncol = 1)
```
## 2023:
```{r echo=FALSE}

plot2023 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2023), aes(x = wy_doy, y = ice_probability)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=215,xmax=243,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Predicted Probability of Ice Over Time",
    x = "Water Year Day of Year",
    y = "Probability of Ice"
  )+ lims(x = c(210,260))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))

plot2023_2 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2023), aes(x = wy_doy, y = cumulative_dis)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=215,xmax=243,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Cumulative Discharge Over Time",
    x = "Water Year Day of Year",
    y = "Cumulative Discharge (cfs)"
  ) + lims(x = c(210,260))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))

grid.arrange(plot2023, plot2023_2, ncol = 1)
```
# Temperature and Probability of Ice:
## 2014:
```{r echo=FALSE}

plot2014_3 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2014), aes(x = wy_doy, y = temperature_C_impute)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=205,xmax=247,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark()+
  labs(
    title = "Temperature Over Time",
    x = "Water Year Day of Year",
    y = "Temperature (C)"
  ) + lims(x = c(200,265))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))
grid.arrange(plot2014,plot2014_3, ncol = 1)
```
## 2015:
```{r echo=FALSE}

plot2015_3 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2015), aes(x = wy_doy, y = temperature_C_impute)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=197,xmax=239,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark()+
  labs(
    title = "Temperature Over Time",
    x = "Water Year Day of Year",
    y = "Temperature (C)"
  ) + lims(x = c(195,265))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))
grid.arrange(plot2015, plot2015_3, ncol = 1)
```
## 2016:
```{r echo=FALSE}

plot2016_3 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2016), aes(x = wy_doy, y = temperature_C_impute)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=195,xmax=251,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark()+
  labs(
    title = "Temperature Over Time",
    x = "Water Year Day of Year",
    y = "Temperature (C)"
  ) + lims(x = c(190,270))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))
grid.arrange(plot2016, plot2016_3, ncol = 1)
```
## 2017:
```{r echo=FALSE}

plot2017_3 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2017), aes(x = wy_doy, y = temperature_C_impute)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=173,xmax=243,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark()+
  labs(
    title = "Temperature Over Time",
    x = "Water Year Day of Year",
    y = "Temperature (C)"
  ) + lims(x = c(170,260))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))
grid.arrange(plot2017, plot2017_3, ncol = 1)
```
## 2018:
```{r echo=FALSE}

plot2018_3 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2018), aes(x = wy_doy, y = temperature_C_impute)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=214,xmax=235,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Temperature Over Time",
    x = "Water Year Day of Year",
    y = "Temperature (C)"
  ) + lims(x = c(200,250))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))
grid.arrange(plot2018, plot2018_3, ncol = 1)
```
## 2019:
```{r echo=FALSE}

plot2019_3 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2019), aes(x = wy_doy, y = temperature_C_impute)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=206,xmax=260,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Temperature Over Time",
    x = "Water Year Day of Year",
    y = "Temperature (C)"
  )+ lims(x = c(200,270)) +
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))
grid.arrange(plot2019, plot2019_3, ncol = 1)
```
## 2020:
```{r echo=FALSE}

plot2020_3 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2020), aes(x = wy_doy, y = temperature_C_impute)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=211,xmax=234,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Temperature Over Time",
    x = "Water Year Day of Year",
    y = "Temperature (C)"
  ) + lims(x = c(200,250))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))
grid.arrange(plot2020, plot2020_3, ncol = 1)
```
## 2021:
```{r echo=FALSE}

plot2021_3 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2021), aes(x = wy_doy, y = temperature_C_impute)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=210,xmax=255,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Temperature Over Time",
    x = "Water Year Day of Year",
    y = "Temperature (C)"
  ) + lims(x = c(200,270))+
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))
grid.arrange(plot2021, plot2021_3, ncol = 1)
```
## 2022:
```{r echo=FALSE}

plot2022_3 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2022), aes(x = wy_doy, y = temperature_C_impute)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=216,xmax=244,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Temperature Over Time",
    x = "Water Year Day of Year",
    y = "Temperature (C)"
  )+ lims(x = c(210,270)) +
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))
grid.arrange(plot2022, plot2022_3, ncol = 1)
```
## 2023:
```{r echo=FALSE}

plot2023_3 <- ggplot(imputed_data_trimmed_prob %>% filter(waterYear == 2023), aes(x = wy_doy, y = temperature_C_impute)) +
  geom_line(color = "blue") +
  geom_rect(aes(xmin=215,xmax=243,ymin=-Inf,ymax=Inf), fill="lightblue",alpha=0.01)+
  theme_dark() +
  labs(
    title = "Temperature Over Time",
    x = "Water Year Day of Year",
    y = "Temperature (C)"
  ) + lims(x = c(210,260))+ 
  theme(panel.ontop=TRUE,panel.background = element_rect(fill = NA))
grid.arrange(plot2023, plot2023_3, ncol = 1)
```
As you can see, there are no temperature data for 2021 - 2023, and there is a large gap in temperature data in 2019. These years will be left out of the function used to determine the threshold of probability to use when there is no ice on The Loch.

# Plot predicted ice-off dates as a line on the plots earlier in this document
## 2014:
```{r}

plot2014 <- plot2014 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2014],
             linetype="dashed")

plot2014_2 <- plot2014_2 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2014],
             linetype="dashed")


plot2014_3 <- plot2014_3 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2014],
             linetype="dashed")

#Use patchwork library to annotate plots and combine -- I find it easier to use than gridExtra or cowplot
WY2014_all <- (plot2014 / plot2014_2 / plot2014_3) + plot_annotation(
  title = 'WY 2014',
  subtitle = str_wrap('Shaded area indicates windows from weekly photographs (full cover to no coverage); dashed line is the predicted ice-off DOY from the logistical regression models', width=50)
)
WY2014_all

#Export file to new plot to a new Figures folder
ggsave("Figures/WY2014_probability_models_temp_cumulQ.png", dpi=600, width=4, height=8, units="in")

```

## 2015:
```{r echo=FALSE}

plot2015 <- plot2015 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2015],
             linetype="dashed")

plot2015_2 <- plot2015_2 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2015],
             linetype="dashed")

plot2015_3 <- plot2015_3 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2015],
             linetype="dashed")

#Use patchwork library to annotate plots and combine -- I find it easier to use than gridExtra or cowplot
WY2015_all <- (plot2015 / plot2015_2 / plot2015_3) + plot_annotation(
  title = 'WY 2015',
  subtitle = str_wrap('Shaded area indicates windows from weekly photographs (full cover to no coverage); dashed line is the predicted ice-off DOY from the logistical regression models', width=50)
)
WY2015_all

#Export file to new plot to a new Figures folder
ggsave("Figures/WY2015_probability_models_temp_cumulQ.png", dpi=600, width=4, height=8, units="in")

```

## 2016:
```{r}
plot2016 <- plot2016 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2016],
             linetype="dashed")

plot2016_2 <- plot2016_2 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2016],
             linetype="dashed")

plot2016_3 <- plot2016_3 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2016],
             linetype="dashed")

#Use patchwork library to annotate plots and combine -- I find it easier to use than gridExtra or cowplot
WY2016_all <- (plot2016 / plot2016_2 / plot2016_3) + plot_annotation(
  title = 'WY 2016',
  subtitle = str_wrap('Shaded area indicates windows from weekly photographs (full cover to no coverage); dashed line is the predicted ice-off DOY from the logistical regression models', width=50)
)
WY2016_all

#Export file to new plot to a new Figures folder
ggsave("Figures/WY2016_probability_models_temp_cumulQ.png", dpi=600, width=4, height=8, units="in")

```
## 2017:
```{r}
plot2017 <- plot2017 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2017],
             linetype="dashed")

plot2017_2 <- plot2017_2 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2017],
             linetype="dashed")

plot2017_3 <- plot2017_3 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2017],
             linetype="dashed")

#Use patchwork library to annotate plots and combine -- I find it easier to use than gridExtra or cowplot
WY2017_all <- (plot2017 / plot2017_2 / plot2017_3) + plot_annotation(
  title = 'WY 2017',
  subtitle = str_wrap('Shaded area indicates windows from weekly photographs (full cover to no coverage); dashed line is the predicted ice-off DOY from the logistical regression models', width=50)
)
WY2017_all

#Export file to new plot to a new Figures folder
ggsave("Figures/WY2017_probability_models_temp_cumulQ.png", dpi=600, width=4, height=8, units="in")

```
## 2018:
```{r}
plot2018 <- plot2018 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2018],
             linetype="dashed")

plot2018_2 <- plot2018_2 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2018],
             linetype="dashed")

plot2018_3 <- plot2018_3 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2018],
             linetype="dashed")

#Use patchwork library to annotate plots and combine -- I find it easier to use than gridExtra or cowplot
WY2018_all <- (plot2018 / plot2018_2 / plot2018_3) + plot_annotation(
  title = 'WY 2018',
  subtitle = str_wrap('Shaded area indicates windows from weekly photographs (full cover to no coverage); dashed line is the predicted ice-off DOY from the logistical regression models', width=50)
)
WY2018_all

#Export file to new plot to a new Figures folder
ggsave("Figures/WY2018_probability_models_temp_cumulQ.png", dpi=600, width=4, height=8, units="in")

```
## 2019:
```{r}
plot2019 <- plot2019 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2019],
             linetype="dashed")

plot2019_2 <- plot2019_2 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2019],
             linetype="dashed")

plot2019_3 <- plot2019_3 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2019],
             linetype="dashed")

#Use patchwork library to annotate plots and combine -- I find it easier to use than gridExtra or cowplot
WY2019_all <- (plot2019 / plot2019_2 / plot2019_3) + plot_annotation(
  title = 'WY 2019',
  subtitle = str_wrap('Shaded area indicates windows from weekly photographs (full cover to no coverage); dashed line is the predicted ice-off DOY from the logistical regression models', width=50)
)
WY2019_all

#Export file to new plot to a new Figures folder
ggsave("Figures/WY2019_probability_models_temp_cumulQ.png", dpi=600, width=4, height=8, units="in")

```
## 2020:
```{r}
plot2020 <- plot2020 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2020],
             linetype="dashed")

plot2020_2 <- plot2020_2 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2020],
             linetype="dashed")

plot2020_3 <- plot2020_3 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2020],
             linetype="dashed")

#Use patchwork library to annotate plots and combine -- I find it easier to use than gridExtra or cowplot
WY2020_all <- (plot2020 / plot2020_2 / plot2020_3) + plot_annotation(
  title = 'WY 2020',
  subtitle = str_wrap('Shaded area indicates windows from weekly photographs (full cover to no coverage); dashed line is the predicted ice-off DOY from the logistical regression models', width=50)
)
WY2020_all

#Export file to new plot to a new Figures folder
ggsave("Figures/WY2020_probability_models_temp_cumulQ.png", dpi=600, width=4, height=8, units="in")

```
## 2021:
```{r}
plot2021 <- plot2021 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2021],
             linetype="dashed")

plot2021_2 <- plot2021_2 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2021],
             linetype="dashed")

plot2021_3 <- plot2021_3 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2021],
             linetype="dashed")

#Use patchwork library to annotate plots and combine -- I find it easier to use than gridExtra or cowplot
WY2021_all <- (plot2021 / plot2021_2 / plot2021_3) + plot_annotation(
  title = 'WY 2021',
  subtitle = str_wrap('Shaded area indicates windows from weekly photographs (full cover to no coverage); dashed line is the predicted ice-off DOY from the logistical regression models', width=50)
)
WY2021_all

#Export file to new plot to a new Figures folder
ggsave("Figures/WY2021_probability_models_temp_cumulQ.png", dpi=600, width=4, height=8, units="in")

```
## 2022:
```{r}
plot2022 <- plot2022 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2022],
             linetype="dashed")

plot2022_2 <- plot2022_2 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2022],
             linetype="dashed")

plot2022_3 <- plot2022_3 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2022],
             linetype="dashed")

#Use patchwork library to annotate plots and combine -- I find it easier to use than gridExtra or cowplot
WY2022_all <- (plot2022 / plot2022_2 / plot2022_3) + plot_annotation(
  title = 'WY 2022',
  subtitle = str_wrap('Shaded area indicates windows from weekly photographs (full cover to no coverage); dashed line is the predicted ice-off DOY from the logistical regression models', width=50)
)
WY2022_all

#Export file to new plot to a new Figures folder
ggsave("Figures/WY2022_probability_models_temp_cumulQ.png", dpi=600, width=4, height=8, units="in")

```
## 2023:
```{r}
plot2023 <- plot2023 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2023],
             linetype="dashed")

plot2023_2 <- plot2023_2 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2023],
             linetype="dashed")

plot2023_3 <- plot2023_3 +
  geom_vline(xintercept=imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy[imputed_ice_off_dates_predicted_obs$waterYear == 2023],
             linetype="dashed")

#Use patchwork library to annotate plots and combine -- I find it easier to use than gridExtra or cowplot
WY2023_all <- (plot2023 / plot2023_2 / plot2023_3) + plot_annotation(
  title = 'WY 2023',
  subtitle = str_wrap('Shaded area indicates windows from weekly photographs (full cover to no coverage); dashed line is the predicted ice-off DOY from the logistical regression models', width=50)
)
WY2023_all

#Export file to new plot to a new Figures folder
ggsave("Figures/WY2023_probability_models_temp_cumulQ.png", dpi=600, width=4, height=8, units="in")

```

# Average Probability in Observed Windows:
## Extracting only observations and predicted probabilities within observed ice windows:
```{r}
# Function to extract ice_probability between wy_doy_100_ice and wy_doy_0_ice, which are the dates of 100% ice and 0% ice obtained from the photos of The Loch. These dates are contained in the DF "obs_ice_melt_windows".
extract_ice_probability <- function(imputed_data_trimmed_prob, obs_ice_melt_windows) {
  
  # Initialize an empty list to store the results
  result_list <- list()
  
  # Loop through each unique waterYear
  for (year in unique(obs_ice_melt_windows$waterYear)) {
    
    # Get the doy range for the current waterYear
    doy_100 <- obs_ice_melt_windows$wy_doy_100_ice[obs_ice_melt_windows$waterYear == year]
    doy_0 <- obs_ice_melt_windows$wy_doy_0_ice[obs_ice_melt_windows$waterYear == year]
    
    # Subset the imputed_data_trimmed_prob data frame for the current year
    prob_data <- imputed_data_trimmed_prob[imputed_data_trimmed_prob$waterYear == year, ]
    
    # Filter the rows where wy_doy is between wy_doy_100_ice and wy_doy_0_ice
    filtered_data <- prob_data[prob_data$wy_doy >= doy_100 & prob_data$wy_doy <= doy_0, ]
    
    # Store the filtered data in the result list
    result_list[[as.character(year)]] <- filtered_data
  }
  
  # Combine all the filtered data frames into one
  probs_df <- do.call(rbind, result_list)
  
  return(probs_df)
}

# Applying the function to our data
probs_df <- extract_ice_probability(imputed_data_trimmed_prob, obs_ice_melt_windows)
# filtering to remove 2021-2023 and 2019:
probs_df <- probs_df %>% filter(waterYear<2021 & waterYear != 2019)
# View the result
#View(probs_df)
```
Now we can calculate the average probability of each year within only the timeframe of observed ice melt.
```{r}
# calculate the average probability by year:
avg_probs_df <- group_by(probs_df, waterYear) %>%
  summarize(average_prob = mean(ice_probability))
# calculate the average probability across all years
avg_prob_all_years <- mean(avg_probs_df$average_prob)
print(avg_prob_all_years)

# calculate the max probability by year:
max_probs_df <- group_by(probs_df, waterYear) %>%
  summarize(max_prob = max(ice_probability))
# calculate the median max probability across all years
max_prob_all_years <- median(max_probs_df$max_prob)
print(max_prob_all_years)
```
## Graph average probability by year (cumul temp):
```{r}
ggplot(avg_probs_df, aes(x = waterYear, y = average_prob)) +
  geom_point(color = "red") +         # Add points for each year
  labs(title = "Average Ice Probability Over Time",
       x = "Year",
       y = "Average Ice Probability")

ggplot(max_probs_df, aes(x = waterYear, y = max_prob)) +
  geom_point(color = "red") +         # Add points for each year
  labs(title = "Max Ice Probability Over Time",
       x = "Year",
       y = "Max Ice Probability")
```
# Accuracy Comparison
Using value of average probability of all years within observed windows for imputed temp as threshold and comparing it to randomly chosen probability threshold of 0.5
## 0.5 as Threshold
```{r}
# create a vector of predicted probabilities
preds_05 <- predict(trimmed_imputed_all,
  newdata = select(imputed_data_trimmed_14_23, -ice_presence), # remove real outcomes
  type = "response"
)

# if probability < threshold, ice IS on The Loch
preds_outcome_05 <- ifelse(preds_05 < 0.5,
  1,
  0
)

# transform predictions into factor and set labels
preds_outcome_05 <- factor(preds_outcome_05,
  levels = c(1, 0),
  labels = c("ice", "no ice")
)

# compare observed vs. predicted outcome
tab_05 <- table(imputed_data_trimmed_14_23$ice_presence, preds_outcome_05,
  dnn = c("observed", "predicted")
)

# print results
tab_05

accuracy_05 <- sum(diag(tab_05)) / sum(tab_05)
accuracy_05

# sensitivity
sensitivity_05 <- tab_05[2, 2] / (tab_05[2, 2] + tab_05[2, 1])
sensitivity_05

# specificity
specificity_05<- tab_05[1, 1] / (tab_05[1, 1] + tab_05[1, 2])
specificity_05
```
## Average Probability from imputed temp as threshold:
```{r}
# create a vector of predicted probabilities
preds_avg <- predict(trimmed_imputed_reduced_final,
  newdata = select(imputed_data_trimmed_14_23, -ice_presence), # remove real outcomes
  type = "response"
)

# if probability < threshold, ice IS on The Loch
preds_outcome_avg <- ifelse(preds_avg < avg_prob_all_years,
  1,
  0
)

# transform predictions into factor and set labels
preds_outcome_avg <- factor(preds_outcome_avg,
  levels = c(1, 0),
  labels = c("ice", "no ice")
)

# compare observed vs. predicted outcome
tab_avg <- table(imputed_data_trimmed_14_23$ice_presence, preds_outcome_avg,
  dnn = c("observed", "predicted")
)

# print results
tab_avg

accuracy_avg <- sum(diag(tab_avg)) / sum(tab_avg)
accuracy_avg

# sensitivity
sensitivity_avg <- tab_avg[2, 2] / (tab_avg[2, 2] + tab_avg[2, 1])
sensitivity_avg

# specificity
specificity_avg<- tab_avg[1, 1] / (tab_avg[1, 1] + tab_avg[1, 2])
specificity_avg
```
After performing these analyses, we can see that the chosen threshold of 0.5 is actually more accurate than the average probability across all years. So we will use 0.5 for our threshold.

# Ice-Off Dates:
Pulling out the first day each year when ice is "no ice" during the time period when we have observed ice-off dates.

First we have to factor the predicted probability of ice to "ice" and "no ice"
```{r}
probability_impute_0_1 <- ifelse(probabilities_impute < 0.5,
  1,
  0
)
# changing the labels of the probabilities
probability_impute_factor <- factor(probability_impute_0_1, levels = c(1, 0),labels = c("ice", "no ice"))


imputed_data_trimmed_prob$predicted_ice <- probability_impute_factor
```

```{r}
# removing years with no data
imputed_data_trimmed_prob_filter <- imputed_data_trimmed_prob %>% filter(waterYear<2021)

# Create an empty data frame to store the results
imputed_ice_off_dates_predicted_obs <- data.frame()

# Iterate through each unique waterYear
for(year in unique(imputed_data_trimmed_prob_filter$waterYear)) {
  
  # Filter the data for the current year
  year_data <- imputed_data_trimmed_prob_filter %>% filter(waterYear == year)
  
  # Find the first day when ice_presence == "no ice"
  first_no_ice_presence <- year_data %>% 
    filter(ice_presence == "no ice") %>%
    arrange(wy_doy) %>%
    slice(1)
  
  # Find the first day when predicted_ice == "no ice"
  first_no_ice_predicted <- year_data %>% 
    filter(predicted_ice == "no ice") %>%
    arrange(wy_doy) %>%
    slice(1)
  
  # Create a new row with the results
  result_row <- data.frame(
    waterYear = year,
    first_no_ice_presence_doy = first_no_ice_presence$wy_doy,
    first_no_ice_predicted_doy = first_no_ice_predicted$wy_doy
  )
  
  # Append the row to the result data frame
  imputed_ice_off_dates_predicted_obs <- bind_rows(imputed_ice_off_dates_predicted_obs, result_row)
}

# find the difference between predicted and observed and print the mean of this difference across all years
imputed_ice_off_dates_predicted_obs$difference <- ((imputed_ice_off_dates_predicted_obs$first_no_ice_presence_doy)-(imputed_ice_off_dates_predicted_obs$first_no_ice_predicted_doy))*-1
print(mean(imputed_ice_off_dates_predicted_obs$difference))
# negative means the predicted ice-ff date was EARLIER than the observed


# View the result data frame
View(imputed_ice_off_dates_predicted_obs)
```
Now that we have a threshold and we saw how our model performed, we can hindcast! Go to Rmd "03_Hindcasting.Rmd".

# Functional Ice-Off Dates:
Pulling out the first day each year when ice is "no ice" during the time period when we have observed ice-off dates.

```{r}
probabilities_impute_func <- predict(trimmed_imputed_reduced_final_func,
  newdata = select(imputed_data_trimmed_14_23_func, -ice_presence), # remove real outcomes
  type = "response")

imputed_data_trimmed_prob_func <- imputed_data_trimmed_14_23_func
imputed_data_trimmed_prob_func$ice_probability <- probabilities_impute_func
```


First we have to factor the predicted probability of ice to "ice" and "no ice"
```{r}
probability_impute_0_1_func <- ifelse(probabilities_impute_func < 0.5,
  1,
  0
)
# changing the labels of the probabilities
probability_impute_factor_func <- factor(probability_impute_0_1_func, levels = c(1, 0),labels = c("ice", "no ice"))


imputed_data_trimmed_prob_func$predicted_ice <- probability_impute_factor_func
```


```{r}
# removing years with no data
imputed_data_trimmed_prob_filter_func <- imputed_data_trimmed_prob_func %>% filter(waterYear<2021)

# Create an empty data frame to store the results
imputed_ice_off_dates_predicted_obs_func <- data.frame()

# Iterate through each unique waterYear
for(year in unique(imputed_data_trimmed_prob_filter_func$waterYear)) {
  
  # Filter the data for the current year
  year_data <- imputed_data_trimmed_prob_filter_func %>% filter(waterYear == year)
  
  # Find the first day when ice_presence == "no ice"
  first_no_ice_presence <- year_data %>% 
    filter(ice_presence == "no ice") %>%
    arrange(wy_doy) %>%
    slice(1)
  
  # Find the first day when predicted_ice == "no ice"
  first_no_ice_predicted <- year_data %>% 
    filter(predicted_ice == "no ice") %>%
    arrange(wy_doy) %>%
    slice(1)
  
  # Create a new row with the results
  result_row <- data.frame(
    waterYear = year,
    first_no_ice_presence_doy = first_no_ice_presence$wy_doy,
    first_no_ice_predicted_doy = first_no_ice_predicted$wy_doy
  )
  
  # Append the row to the result data frame
  imputed_ice_off_dates_predicted_obs_func <- bind_rows(imputed_ice_off_dates_predicted_obs_func, result_row)
}

# find the difference between predicted and observed and print the mean of this difference across all years
imputed_ice_off_dates_predicted_obs_func$difference <- ((imputed_ice_off_dates_predicted_obs_func$first_no_ice_presence_doy)-(imputed_ice_off_dates_predicted_obs_func$first_no_ice_predicted_doy))*-1
print(mean(imputed_ice_off_dates_predicted_obs_func$difference))
# negative means the predicted ice-ff date was EARLIER than the observed


# View the result data frame
View(imputed_ice_off_dates_predicted_obs_func)
```



## WY2016-2018 plot
```{r}
WY2016_2018 <- (WY2016_all | WY2017_all | WY2018_all) + plot_annotation(
  title = 'WY 2016-2018',
  subtitle = str_wrap('Shaded area indicates windows from weekly photographs (full cover to no coverage); dashed line is the predicted ice-off DOY from the logistical regression models', width=110)
)
WY2016_2018

#Export file to new plot to a new Figures folder
ggsave("Figures/WY2016_2018_probability_models_temp_cumulQ.png", dpi=600, width=12, height=10, units="in")
```
```{r}
WY2016_noProb <- (plot2016_2 / plot2016_3)
WY2017_noProb <- (plot2017_2 / plot2017_3)
WY2018_noProb <- (plot2018_2 / plot2018_3)

WY2016_2018_noProb <- (WY2016_noProb | WY2017_noProb | WY2018_noProb) + plot_annotation(
  title = 'WY 2016-2018',
  subtitle = str_wrap('Shaded area indicates windows from weekly photographs (full cover to no coverage); dashed line is the predicted ice-off DOY from the logistical regression models', width=110)
)
WY2016_2018_noProb

#Export file to new plot to a new Figures folder
ggsave("Figures/WY2016_2018_temp_cond.png", dpi=600, width=12, height=10, units="in")
```
