---
title: "Hindcasting"
output: html_document
date: "2024-11-23"
---

```{r}
# load libraries
library(pacman)
p_load(dplyr,dataRetrieval,lubridate,tidyr,ggplot2,viridis,readxl,imputeTS,tsibble,sjPlot,pROC,gridExtra,broom,gtsummary)
# This is the file that combines and cleans data:
source("Input_Files/01_Data_Input.R")
```

# Hindcasting:
```{r}
hindcast_preds <- predict(trimmed_imputed_reduced_final,
  newdata = imputed_data_trimmed,
  type = "response")

# making the probability a binary using the threshold derived from the average of daily temp model probabilities
hindcast_probs_0_1 <- ifelse(hindcast_preds < 0.5,
  1,
  0
)
# changing the labels of the probabilities
hindcast_prob_factor <- factor(hindcast_probs_0_1, levels = c(1, 0),labels = c("ice", "no ice"))

# adding the probabilities to the main df
hindcast_imputed_df <- imputed_data_trimmed  
hindcast_imputed_df$predicted_ice <- hindcast_prob_factor
```

Pulling out the first day each year when ice is "no ice" for hindcasted data
```{r}
# Initialize an empty data frame to store the results
hindcasted_ice_off_dates <- data.frame()

# Iterate through each unique waterYear
for(year in unique(hindcast_imputed_df$waterYear)) {
  
  # Filter the data for the current year where predicted_ice == "no ice"
  year_data <- hindcast_imputed_df %>% 
    filter(waterYear == year, predicted_ice == "no ice") %>%
    arrange(wy_doy)  # Sort by wy_doy to find the first day
  
  # If there is any day where predicted_ice == "no ice"
  if (nrow(year_data) > 0) {
    # Get the first day (earliest day) in that year where predicted_ice == "no ice"
    first_no_ice_day <- year_data %>% slice(1)
    
    # Create a new row with the waterYear and first wy_doy
    result_row <- data.frame(
      waterYear = year,
      first_no_ice_doy = first_no_ice_day$wy_doy
    )
    
    # Append the result_row to the result_df
    hindcasted_ice_off_dates <- bind_rows(hindcasted_ice_off_dates, result_row)
  }
}

# View the result data frame
print(hindcasted_ice_off_dates)
```
Plot them!
```{r}
hindcast_plot <- ggplot(hindcasted_ice_off_dates %>% filter(waterYear<2021), aes(x = waterYear, y = first_no_ice_doy)) +
  geom_point() +
  geom_smooth()+
  scale_x_continuous(
    breaks = seq(min(hindcasted_ice_off_dates$waterYear), max(hindcasted_ice_off_dates$waterYear), by = 1))+
  labs(
    title = "Hindcasted Date of Ice-Off 1984-2023",
    x = "Year",
    y = "Water Year DOY of Ice-Off")+
   theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))
hindcast_plot
```
As we saw in the analysis, 2021-2023 have a large amount of missing temperature values, so I exclude them from this plot until we can track down the missing temperature observations.










# Hindcasting (Functional Ice-Off):
```{r}
hindcast_preds_func <- predict(trimmed_imputed_reduced_final_func,
  newdata = imputed_data_trimmed_func,
  type = "response")

# making the probability a binary using the threshold derived from the average of daily temp model probabilities
hindcast_probs_0_1_func <- ifelse(hindcast_preds_func < 0.5,
  1,
  0
)
# changing the labels of the probabilities
hindcast_prob_factor_func <- factor(hindcast_probs_0_1_func, levels = c(1, 0),labels = c("ice", "no ice"))

# adding the probabilities to the main df
hindcast_imputed_df_func <- imputed_data_trimmed_func 
hindcast_imputed_df_func$predicted_ice <- hindcast_prob_factor_func
```

Pulling out the first day each year when ice is "no ice" for hindcasted data
```{r}
# Initialize an empty data frame to store the results
hindcasted_ice_off_dates_func <- data.frame()

# Iterate through each unique waterYear
for(year in unique(hindcast_imputed_df_func$waterYear)) {
  
  # Filter the data for the current year where predicted_ice == "no ice"
  year_data <- hindcast_imputed_df_func %>% 
    filter(waterYear == year, predicted_ice == "no ice") %>%
    arrange(wy_doy)  # Sort by wy_doy to find the first day
  
  # If there is any day where predicted_ice == "no ice"
  if (nrow(year_data) > 0) {
    # Get the first day (earliest day) in that year where predicted_ice == "no ice"
    first_no_ice_day <- year_data %>% slice(1)
    
    # Create a new row with the waterYear and first wy_doy
    result_row_func <- data.frame(
      waterYear = year,
      first_no_ice_doy = first_no_ice_day$wy_doy
    )
    
    # Append the result_row to the result_df
    hindcasted_ice_off_dates_func <- bind_rows(hindcasted_ice_off_dates_func, result_row_func)
  }
}

# View the result data frame
print(hindcasted_ice_off_dates_func)
```
Plot them!
```{r}
hindcast_plot_func <- ggplot(hindcasted_ice_off_dates_func %>% filter(waterYear<2021), aes(x = waterYear, y = first_no_ice_doy)) +
  geom_point() +
  geom_smooth()+
  scale_x_continuous(
    breaks = seq(min(hindcasted_ice_off_dates_func$waterYear), max(hindcasted_ice_off_dates_func$waterYear), by = 1))+
  labs(
    title = "Hindcasted Date of Ice-Off 1984-2023",
    x = "Year",
    y = "Water Year DOY of Ice-Off")+
   theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))
hindcast_plot_func
```