---
title: "20240925_Model_Building"
output: html_document
date: "2024-09-25"
---
# Libraries and Functions
```{r}
library(pacman)
p_load(dplyr,dataRetrieval,lubridate,tidyr,ggplot2,viridis,readxl,imputeTS,tsibble,sjPlot,pROC)

source("Input_Files/00_functions.R")
```
# Outlet - Bring in Cumulative Flow
```{r}
# Get data for LV
lv_no <- '401733105392404'

# define parameters of interest, and get those parameter names
params <- c('00060', '00671', '80154', '00665')

# get daily values from NWIS
lv_dat <- readNWISdv(siteNumbers = lv_no, parameterCd = params,
                         startDate = '1983-10-01', endDate = '2024-09-30')

# rename columns using renameNWISColumns from package dataRetrieval
lv_dat <- renameNWISColumns(lv_dat,
                                p00665 = "TP_mgL",
                                p00671 = "Orthophosphate_mgL",
                                p80154 = "SS_mgL")

# use function `grep` to identify which columns are code columns
lv_names <- names(lv_dat)
grep('_cd', lv_names) # returns the index of the match

grep('_cd', lv_names, value = TRUE) # returns the matched elements themselves

# change the code column names to be more explicit about what they contain
# using function gsub
gsub('_cd', '_code', lv_names)

lv_dat <- select(lv_dat, -contains('_cd'))
head(lv_dat)

lv_dat <- addWaterYear(lv_dat)

# calculate cumulative discharge for each year by first grouping by water year,
# and then using the "cumsum" function. Add day of water year for plotting purposes.
# These steps will build a new dataframe, with the existing information in yahara_dat
# but with two additional columns.
cumulative_dat <- group_by(lv_dat, waterYear) %>%
  mutate(cumulative_dis = cumsum(Flow), 
         wy_doy = seq(1:n()))

# ungroup cumulative_dat 
cumulative_dat_ungroup <- cumulative_dat %>%
  ungroup() %>%
  as.data.frame()

# make sure it looks good
str(cumulative_dat_ungroup)

# rename the df so its nicer, remove the site number column
cumulative_flow_df <- cumulative_dat_ungroup %>% select(-site_no) %>% mutate(Date = as.Date(Date, tz = "MST", format = "%Y-%m-%d"))
view(cumulative_flow_df)
```
# Outlet - Bring in Temp and Conductivity Data from Outlet
```{r}
out_cond_dat <- read.csv("Input_Files/Loch_O_daily_conductivity.csv")
out_temp_dat <- read.csv("Input_Files/Loch_O_daily_temperature.csv")
out_condTemp_dat19_23 <- left_join(out_cond_dat,out_temp_dat, by = "Date")
out_condTemp_dat11_19 <- read.csv("Input_Files/LochDaily_TempCond_2011-2019.csv")
View(out_condTemp_dat11_19)
View(out_condTemp_dat)

# merge 2011-2019 with 2019-2023
out_condTemp_allDates <- merge(out_condTemp_dat11_19, out_condTemp_dat19_23, all = TRUE)
View(out_condTemp_allDates)
```
# Bring in Ice off and 20th Percentile dates
```{r}
# ice off dates and 20th percentile dates
ice_off_20thPer_dates <- read_xlsx("Input_Files/20240925_IceOff_20thQuartile_Dates.xlsx")
view(ice_off_20thPer_dates)
# 2013 - 2023 daily ice on (1) or off (0)
ice_off_binary <- read.csv("Input_Files/binary_iceOff_20241001.csv") %>% select(c(Date,ice.0.1.,wy_doy)) %>% rename(ice_or_no=ice.0.1.)
View(ice_off_binary)

# recode ice_or_no into 2 classes
ice_off_binary$ice_presence <- ifelse(ice_off_binary$ice_or_no == 0,
  0,
  1
)
# set labels for ice
ice_off_binary$ice_presence <- factor(ice_off_binary$ice_presence,
  levels = c(1, 0),
  labels = c("ice", "no ice")
)
View(ice_off_binary)
str(ice_off_binary$ice_presence)
levels(ice_off_binary$ice_presence)

# combine binary ice on or off with temp and conductivity
out_dat_and_ice <- left_join(out_condTemp_allDates,ice_off_binary, by = "Date") %>% mutate(Date = as.Date(Date, tz = "MST", format = "%Y-%m-%d"))
View(out_dat_and_ice)

# full_data <- ice_off_binary %>%
#   mutate(Date = ymd(Date)) %>%
#   left_join(., out_condTemp_long) %>%
#   drop_na(temp_or_cond)

```


# Combine all data
```{r}
cumul_out_ice_ALL <- left_join(out_dat_and_ice,cumulative_flow_df,by="Date") %>% select(-wy_doy.y) %>% rename(wy_doy = wy_doy.x) %>% drop_na(ice_or_no)
View(cumul_out_ice_ALL)

ALL_2020 <- cumul_out_ice_ALL %>% filter(waterYear == 2020)
ALL_2021 <- cumul_out_ice_ALL %>% filter(waterYear == 2021)
some_2021 <- cumul_out_ice_ALL %>% filter(waterYear == 2021) %>% filter(wy_doy > 183 & wy_doy <263)
ALL_some <- cumul_out_ice_ALL %>% filter(wy_doy > 183 & wy_doy <263)
view(ALL_some)
```

# Regression
```{r}
model1 <- glm(ice_or_no~Temperature_C+cond_uScm, data = cumul_out_ice_ALL, family = binomial)
summary(model1)

model1a <- glm(ice_or_no~Temperature_C+cond_uScm+Flow, data = cumul_out_ice_ALL, family = binomial)
summary(model1a)

model3 <- lm(ice_or_no~Temperature_C+cond_uScm, data = cumul_out_ice_ALL)
summary(model3)

model4 <- lm(ice_or_no~Temperature_C+cond_uScm+Flow, data = cumul_out_ice_ALL)
summary(model4)

model5 <- lm(ice_or_no~Temperature_C+cond_uScm+cumulative_dis, data = cumul_out_ice_ALL)
summary(model5)

plot(model1a)

model_window <- glm(ice_or_no~Temperature_C+cond_uScm+Flow, data = ALL_some, family = binomial)
summary(model_window)
```
# Binary Regressions following the Tutorial
```{r}
model_allDates <- glm(ice_presence~Temperature_C+cond_uScm+Flow, data = cumul_out_ice_ALL, family = binomial)
summary(model_allDates)

model1Var <- glm(ice_presence~Temperature_C, data = cumul_out_ice_ALL, family = binomial)
summary(model1Var)
exp(coef(model1Var)["Temperature_C"])
# the above exponential function of the coefficient for temperature means that one unit of increase with temperature increases the chances of there being no ice by a factor of 2.424
# the probability of there being no ice increases by 142% when temperature increases by 1

model1Var2021 <- glm(ic_presencee~Temperature_C, data = ALL_2021, family = binomial)
summary(model1Var2021)
exp(coef(model1Var2021)["Temperature_C"])
# in 2021, the probability of there being no ice increases by a factor of 117% when temperature increases by 1
```
# Temp Model Plot, Following Tutorial
```{r}
plot_model(model1Var,
  type = "pred",
  terms = "Temperature_C"
) +
  labs(y = "Prob (no ice)")
```
# Conductivity Model and Plot
```{r}
cond_model <- glm(ice_presence~cond_uScm, data = cumul_out_ice_ALL, family = binomial)
summary(cond_model)
exp(coef(cond_model)["cond_uScm"])
# -29% - probability of no ice decreases by 29% when conductivity rises by one unit (high conductivity = ice)
plot_model(cond_model,
  type = "pred",
  terms = "cond_uScm"
) +
  labs(y = "Prob (no ice)")
```
# Flow model and Plot
```{r}
flow_model <- glm(ice_presence~Flow, data = cumul_out_ice_ALL, family = binomial)
summary(flow_model)
exp(coef(flow_model)["Flow"])
# 21% - probability of no ice increases by 21% when Flow increases by one unit - high flow = no ice
plot_model(flow_model,
  type = "pred",
  terms = "Flow"
) +
  labs(y = "Prob (no ice)")
```
# Cumulative Flow Model and Plot
```{r}
cumul_flow_model <- glm(ice_presence~cumulative_dis, data = cumul_out_ice_ALL, family = binomial)
summary(cumul_flow_model)
exp(coef(cumul_flow_model)["cumulative_dis"])
# probability of no ice increases by < 1% when cumulative flow increases by one unit (cumulative flow units don't show relationship well)
plot_model(cumul_flow_model,
  type = "pred",
  terms = "cumulative_dis"
) +
  labs(y = "Prob (no ice)")
```
# All factors?
```{r}
all_model <- glm(ice_presence~Flow+Temperature_C+cond_uScm+waterYear, data = cumul_out_ice_ALL, family = binomial)
summary(all_model)

reduced_model <- glm(ice_presence~Temperature_C+cond_uScm+waterYear, data = cumul_out_ice_ALL, family = binomial)
# when just flow and water year - relationship between flow and ice is correct. 
# when flow and temp and water year or flow and cond and water year or flow and temp and cond or all three, relationship between flow and ice is wrong...
anova(reduced_model, all_model,
  test = "LRT"
)

# OR and 95% CI
round(exp(cbind(OR = coef(reduced_model), confint(reduced_model))), 3)
# only the ORs with significant p-values
round(exp(coef(reduced_model))[coef(summary(reduced_model))[, "Pr(>|z|)"] < 0.05], 3)
# probability of no ice decreases with increase in flow and conductivity, and increases with increase in temperature and water year -- high FLOW(?) and conductivity = ice, high temp and later year = no ice

plot_model(reduced_model,
  type = "pred",
  terms = c("Temperature_C [all]","waterYear [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")

plot_model(reduced_model,
  type = "pred",
  terms = c("cond_uScm [all]","waterYear [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")
```
# Testing interaction
```{r}
inter_model <- glm(ice_presence~Temperature_C*cond_uScm*Flow*waterYear, data = cumul_out_ice_ALL, family = binomial)
summary(inter_model)
anova(reduced_model, inter_model,
  test = "LRT"
)
round(exp(coef(inter_model))[coef(summary(inter_model))[, "Pr(>|z|)"] < 0.05], 3)

inter_model2 <- glm(ice_presence~Temperature_C+cond_uScm*Flow*waterYear, data = cumul_out_ice_ALL, family = binomial)
summary(inter_model2)

round(exp(coef(inter_model2))[coef(summary(inter_model2))[, "Pr(>|z|)"] < 0.05], 3)

anova(inter_model2, inter_model,
  test = "LRT"
)

inter_model3 <- glm(ice_presence~Temperature_C+cond_uScm*Flow+waterYear, data = cumul_out_ice_ALL, family = binomial)
summary(inter_model3)

round(exp(coef(inter_model3))[coef(summary(inter_model3))[, "Pr(>|z|)"] < 0.05], 3)

anova(inter_model3, inter_model2,
  test = "LRT"
)

plot_model(inter_model3,
  type = "pred",
  terms = c("Flow [all]","waterYear [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")
# what the heck is this

inter_model4 <- glm(ice_presence~Temperature_C*cond_uScm*waterYear, data = cumul_out_ice_ALL, family = binomial)
summary(inter_model4)

round(exp(coef(inter_model4))[coef(summary(inter_model4))[, "Pr(>|z|)"] < 0.05], 3)

anova(inter_model4, inter_model3,
  test = "LRT"
)

inter_model5 <- glm(ice_presence~Temperature_C+cond_uScm*waterYear, data = cumul_out_ice_ALL, family = binomial)
summary(inter_model5)

round(exp(coef(inter_model5))[coef(summary(inter_model5))[, "Pr(>|z|)"] < 0.05], 3)

anova(inter_model5, inter_model4,
  test = "LRT")

plot_model(inter_model5,
  type = "pred",
  terms = c("Temperature_C [all]","waterYear [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")

plot_model(inter_model5,
  type = "pred",
  terms = c("cond_uScm","waterYear [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")

```
# Model selection
```{r}
# save initial model
all_variables_m <- glm(ice_presence ~ (Flow+cumulative_dis+Temperature_C+cond_uScm+waterYear)^2,
  data = cumul_out_ice_ALL,
  family = "binomial"
)
# select best model according to AIC using mixed selection
AIC_chosen_model <- step(all_variables_m,
  direction = "both", # both = mixed selection
  trace = FALSE # do not display intermediate steps
)
# display results of final model
summary(AIC_chosen_model)
```
# Testing Predictions
```{r}
# create a vector of predicted probabilities
preds <- predict(AIC_chosen_model,
  newdata = select(cumul_out_ice_ALL, -ice_presence), # remove real outcomes
  type = "response"
)

# if probability < threshold, ice IS on The Loch
preds_outcome <- ifelse(preds < 0.5,
  1,
  0
)

# transform predictions into factor and set labels
preds_outcome <- factor(preds_outcome,
  levels = c(1, 0),
  labels = c("ice", "no ice")
)

# compare observed vs. predicted outcome
tab <- table(cumul_out_ice_ALL$ice_presence, preds_outcome,
  dnn = c("observed", "predicted")
)

# print results
tab

accuracy <- sum(diag(tab)) / sum(tab)
accuracy

# sensitivity
sensitivity <- tab[2, 2] / (tab[2, 2] + tab[2, 1])
sensitivity

# specificity
specificity <- tab[1, 1] / (tab[1, 1] + tab[1, 2])
specificity
```
# AUC and ROC
```{r}
# save roc object
res <- roc(ice_presence ~ fitted(AIC_chosen_model),
  data = cumul_out_ice_ALL
)

# plot ROC curve
ggroc(res, legacy.axes = TRUE)
```


```{r}
pred_values <- predict(model_window, type = "response")
df <- data.frame(observed = ALL_some$ice_or_no, predicted = pred_values)
ggplot(df, aes(y = predicted, x = observed))+
  geom_point()
```



```{r}
model2020 <- glm(ice_or_no~Temperature_C+cond_uScm+Flow, data = ALL_2020, family = binomial)
summary(model2020)
model2021 <- glm(ice_or_no~Temperature_C+cond_uScm+Flow, data = ALL_2021, family = binomial)
summary(model2021)
plot(x = model2021$fitted.values, ALL_2021$ice_or_no)
model2021some <- glm(ice_or_no~Temperature_C+cond_uScm+Flow, data = some_2021, family = binomial)
summary(model2021some)

```


# Long format Temp and Conductivity
```{r}
out_condTemp_long <- pivot_longer(out_condTemp_dat, -Date, names_to="hydro_variables", values_to = "temp_or_cond") %>% mutate(waterYear = calcWaterYear(Date)) %>% mutate(Date = as.Date(Date, tz = "MST", format = "%Y-%m-%d"))%>% mutate(wy_doy = hydro.day(Date))

# write_xlsx(out_condTemp_long ,"Input_Files/20240925_IceOff_20thQuartile_Dates.xlsx")


view(out_condTemp_long)

```


# Plot temp and conductivity by year
## 2020
```{r}
ggplot(out_condTemp_long %>% filter(waterYear==2020), aes(wy_doy,temp_or_cond, group = waterYear, col = hydro_variables)) + geom_point(size = 0.5) + geom_vline(xintercept = 234) # ice off date
```
## 2021
```{r}
ggplot(out_condTemp_long %>% filter(waterYear==2021), aes(wy_doy,temp_or_cond, group = waterYear, col = hydro_variables)) + geom_point(size = 0.5) + geom_vline(xintercept = 254) # ice off date
```
## 2022
```{r}
ggplot(out_condTemp_long %>% filter(waterYear==2022), aes(wy_doy,temp_or_cond, group = waterYear, col = hydro_variables)) + geom_point(size = 0.5) + geom_vline(xintercept = 243) # ice off date
```
## 2023
```{r}
ggplot(out_condTemp_long %>% filter(waterYear==2023), aes(wy_doy,temp_or_cond, group = waterYear, col = hydro_variables)) + geom_point(size = 0.5) + geom_vline(xintercept = 242) # ice off date
```


