---
title: "20240925_Model_Building"
output: html_document
date: "2024-09-25"
---
# Libraries and Functions
```{r}
library(pacman)
p_load(dplyr,dataRetrieval,lubridate,tidyr,ggplot2,viridis,readxl,imputeTS,tsibble,sjPlot,pROC,gridExtra,broom,gtsummary)

source("Input_Files/00_functions.R")

```
# Outlet - Bring in Cumulative Flow
```{r}
# LV Site Number:
# lv_no <- '401733105392404'
# 
# # define parameter for discharge (00060)
# param <- '00060'
# 
# # get daily values from NWIS
# lv_dat <- readNWISdv(siteNumbers = lv_no, parameterCd = param,
#                      startDate = '1983-10-01', endDate = '2024-09-30')
# 
# # rename columns using renameNWISColumns from package dataRetrieval
# # this renames the column for Flow from the parameter ID to "Flow"
# lv_dat <- renameNWISColumns(lv_dat)
# 
# # Removing column with USGS code for observations
# lv_dat <- select(lv_dat, -contains('_cd'))
# 
# # Adding the water year to the df
# lv_dat <- addWaterYear(lv_dat)
# 
# # calculating cumulative discharge for each year by first grouping by water year,
# # and then using the "cumsum" function. Add day of water year for plotting purposes.
# cumulative_dat <- group_by(lv_dat, waterYear) %>%
#   mutate(cumulative_dis = cumsum(Flow), 
#          wy_doy = seq(1:n()))
# 
# # ungroup the dataframe
# cumulative_dat_ungroup <- cumulative_dat %>%
#   ungroup() %>%
#   as.data.frame()
# 
# # rename the df, remove the site number column, make sure dates are in date format. This is the final df
# cumulative_flow_df <- cumulative_dat_ungroup %>% select(-site_no) %>% mutate(Date = as.Date(Date, tz = "MST", format = "%Y-%m-%d"))
# View(cumulative_flow_df)
```
# Outlet Weekly Temp and Cond:
```{r}
# parameterCd <- c('00095','00010')
# lv_dat_outlet_input <- readWQPqw(paste0("USGS-", lv_no), parameterCd)

#Clean up dataframe
# lv_dat_outlet <- lv_dat_outlet_input %>%
#   select(ActivityStartDate, ActivityConductingOrganizationText, CharacteristicName, ResultMeasureValue) %>%
#   pivot_wider(names_from = CharacteristicName, values_from = ResultMeasureValue, values_fn = mean) %>%
#   rename(cond_uScm="Specific conductance",
#          temperature_C_raw="Temperature, water",
#          Date = "ActivityStartDate") %>%
#   select(-ActivityConductingOrganizationText) %>%
#   mutate(wy_doy = hydro.day(Date)) %>%
#   addWaterYear() %>% 
#   distinct(Date, .keep_all = TRUE) %>%
#   as_tsibble(., key = waterYear, index = Date) %>% #time series tibble
#   fill_gaps() #%>%  #makes the missing data implicit
# View(lv_dat_outlet) # THIS IS WEEKLY
# # Imputed data:
# lv_out_impute <- lv_dat_outlet %>%
#   mutate(cond_uScm_impute = imputeTS::na_interpolation(cond_uScm, maxgap = 7),
#        temperature_C_impute = imputeTS::na_interpolation(temperature_C_raw, maxgap = 7))
# 
# # Binding the weekly data and imputed data with cumulative flow
# flow_temp_cond_weekly <- full_join(cumulative_flow_df, lv_dat_outlet, by = "Date")
# flow_temp_cond_impute <- full_join(cumulative_flow_df, lv_out_impute, by = "Date")
# View(flow_temp_cond_impute)
# now sort them for 2011-2023 and bind them with the ice_presence by making a new data frame with select(ice_presence, Date) from the big data set


# Joining weekly and imputed data with the ice presence
# weekly data frame:
# flow_temp_cond_weekly_ice <- full_join(flow_temp_cond_weekly, ice_presence_df, by = "Date") %>% select(Date, wy_doy.x, waterYear.x, Flow, cumulative_dis, cond_uScm, temperature_C_raw, ice_or_no,ice_presence) %>% rename(wy_doy=wy_doy.x,waterYear=waterYear.x)
# View(flow_temp_cond_weekly_ice)

# # imputed data frame:
# flow_temp_cond_imputed_ice <- full_join(flow_temp_cond_impute, ice_presence_df, by = "Date") %>% select(Date, wy_doy.x, waterYear.x, Flow, cumulative_dis, cond_uScm_impute, temperature_C_impute, ice_or_no,ice_presence) %>% rename(wy_doy=wy_doy.x,waterYear=waterYear.x)
# View(flow_temp_cond_imputed_ice)

# # trimming the data frames for windows:
# imputed_data_trimmed <- filter_by_year_and_doy(flow_temp_cond_imputed_ice, c(183,288)) # April 1 - July 15
# weekly_data_trimmed <- filter_by_year_and_doy(flow_temp_cond_weekly_ice, c(183,288)) # April 1 - July 15
```
 
# Plotting weekly and imputed data
```{r}
p1 <- ggplot(out_cond_temp_daily %>% filter(waterYear==2012)%>% filter(wy_doy >188 & wy_doy < 288), aes(x = wy_doy)) +
  geom_point(aes(y = Temperature_C, color = "Daily Temperature (C) 2012")) +
  geom_vline(xintercept = 207)+
  labs(color = "Legend") +
  theme_minimal() +
  theme(legend.position = "top")+
  lims(y = c(0, 11.5))

p2 <- ggplot(flow_temp_cond_impute %>% filter(waterYear.x==2012)%>% filter(wy_doy.x >188 & wy_doy.x < 288), aes(x = wy_doy.x)) +
  geom_point(aes(y = temperature_C_impute, color = "Imputed Temperature (C) 2012")) +
  geom_vline(xintercept = 207)+
  labs(color = "Legend") +
  theme_minimal() +
  theme(legend.position = "top")+
  lims(y = c(0, 11.5))

grid.arrange(p1, p2, ncol = 2)
```









## Imputed Modeling
```{r}
imputeModel <- glm(ice_presence~cumulative_dis+temperature_C_impute, data = imputed_data_trimmed, family = binomial)
summary(imputeModel)
```
## Accuracy of Imputed
```{r}
# create a vector of predicted probabilities
predsImpute <- predict(imputeModel,
  newdata = select(imputed_data_trimmed, -ice_presence), # remove real outcomes
  type = "response"
)

# if probability < threshold, ice IS on The Loch
preds_outcomeImpute <- ifelse(predsImpute < 0.5,
  1,
  0
)

# transform predictions into factor and set labels
preds_outcomeImpute <- factor(preds_outcomeImpute,
  levels = c(1, 0),
  labels = c("ice", "no ice")
)

# compare observed vs. predicted outcome
tabImpute <- table(imputed_data_trimmed$ice_presence, preds_outcomeImpute,
  dnn = c("observed", "predicted")
)

# print results
tabImpute

accuracyImpute <- sum(diag(tabImpute)) / sum(tabImpute)
accuracyImpute

# sensitivity
sensitivityImpute <- tabImpute[2, 2] / (tabImpute[2, 2] + tabImpute[2, 1])
sensitivityImpute

# specificity
specificityImpute <- tabImpute[1, 1] / (tabImpute[1, 1] + tabImpute[1, 2])
specificityImpute
```
## Imputed Plots
```{r}
plot_model(imputeModel,
  type = "pred",
  terms = c("cumulative_dis [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")

plot_model(imputeModel,
  type = "pred",
  terms = c("temperature_C_impute [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")
```



## Weekly Modeling:
```{r}
weeklyModel <- glm(ice_presence~cumulative_dis+temperature_C_raw, data = weekly_data_trimmed, family = binomial)
summary(weeklyModel)
```
## Weekly Accuracy:
```{r}
# create a vector of predicted probabilities
predsWeekly <- predict(weeklyModel,
  newdata = select(weekly_data_trimmed, -ice_presence), # remove real outcomes
  type = "response"
)

# if probability < threshold, ice IS on The Loch
preds_outcomeWeekly <- ifelse(predsWeekly < 0.5,
  1,
  0
)

# transform predictions into factor and set labels
preds_outcomeWeekly <- factor(preds_outcomeWeekly,
  levels = c(1, 0),
  labels = c("ice", "no ice")
)

# compare observed vs. predicted outcome
tabWeekly <- table(weekly_data_trimmed$ice_presence, preds_outcomeWeekly,
  dnn = c("observed", "predicted")
)

# print results
tabWeekly

accuracyWeekly <- sum(diag(tabWeekly)) / sum(tabWeekly)
accuracyWeekly

# sensitivity
sensitivityWeekly <- tabWeekly[2, 2] / (tabWeekly[2, 2] + tabWeekly[2, 1])
sensitivityWeekly

# specificity
specificityWeekly <- tabWeekly[1, 1] / (tabWeekly[1, 1] + tabWeekly[1, 2])
specificityWeekly
```
## Weekly Plots
```{r}
plot_model(weeklyModel,
  type = "pred",
  terms = c("cumulative_dis [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")

plot_model(weeklyModel,
  type = "pred",
  terms = c("temperature_C_raw [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")
```
# Comparing Models
```{r}
tab_model(weeklyModel, imputeModel, reducedTrimModel,
  show.aic = TRUE, # display AIC
  p.style = "numeric_stars" # display p-values and stars
)
```



# Outlet - Bring in Temp and Conductivity Data from Outlet
```{r}
# out_cond_dat <- read.csv("Input_Files/Loch_O_daily_conductivity.csv")
# out_temp_dat <- read.csv("Input_Files/Loch_O_daily_temperature.csv")
# out_condTemp_dat19_23 <- full_join(out_cond_dat,out_temp_dat, by = "Date")
# out_condTemp_dat11_19 <- read.csv("Input_Files/LochDaily_TempCond_2011-2019.csv")
# View(out_condTemp_dat11_19)
# 
# # merge 2011-2019 with 2019-2023
# out_condTemp_allDates <- merge(out_condTemp_dat11_19, out_condTemp_dat19_23, all = TRUE)
# View(out_condTemp_allDates)
# 
# # making a df for daily temp and cond for plotting
# out_cond_temp_daily <- out_condTemp_allDates %>% mutate(Date = as.Date(Date, tz = "MST", format = "%Y-%m-%d")) %>% mutate(wy_doy = hydro.day(Date)) %>% addWaterYear() %>% 
#   distinct(Date, .keep_all = TRUE)
# View(out_cond_temp_daily)

```
# Bring in Ice off and 20th Percentile dates
```{r}
# # ice off dates and 20th percentile dates
# ice_off_20thPer_dates <- read_xlsx("Input_Files/20240925_IceOff_20thQuartile_Dates.xlsx")
# View(ice_off_20thPer_dates)
# # 2013 - 2023 daily ice on (1) or off (0)
# ice_off_binary <- read.csv("Input_Files/binary_iceOff_20241001.csv") %>% select(c(Date,ice.0.1.,wy_doy)) %>% rename(ice_or_no=ice.0.1.) %>% mutate(Date = as.Date(Date, tz = "MST", format = "%Y-%m-%d"))
# View(ice_off_binary)
# 
# # recode ice_or_no into 2 classes
# ice_off_binary$ice_presence <- ifelse(ice_off_binary$ice_or_no == 0,
#   0,
#   1
# )
# # set labels for ice
# ice_off_binary$ice_presence <- factor(ice_off_binary$ice_presence,
#   levels = c(1, 0),
#   labels = c("ice", "no ice")
# )
# View(ice_off_binary)
# str(ice_off_binary$ice_presence)
# levels(ice_off_binary$ice_presence)

# making a df for ice_presence and Date to join to these bad boys
# ice_presence_df <- ice_off_binary %>% select(Date, ice_or_no, ice_presence)
# 
# # combine binary ice on or off with DAILY temp and conductivity
# out_dat_and_ice_daily <- full_join(out_cond_temp_daily,ice_presence_df, by = "Date")
# View(out_dat_and_ice_daily)

# full_data <- ice_off_binary %>%
#   mutate(Date = ymd(Date)) %>%
#   left_join(., out_condTemp_long) %>%
#   drop_na(temp_or_cond)

```


# Combine all data
```{r}
# flow_temp_cond_daily_ice <- left_join(out_dat_and_ice_daily,cumulative_flow_df,by="Date") %>% select(-c(wy_doy.y,waterYear.y)) %>% rename(c(wy_doy = wy_doy.x,waterYear = waterYear.x)) %>% drop_na(ice_or_no)
# View(flow_temp_cond_daily_ice)

ALL_2020 <- flow_temp_cond_daily_ice %>% filter(waterYear == 2020)
ALL_2021 <- flow_temp_cond_daily_ice %>% filter(waterYear == 2021)
some_2021 <- flow_temp_cond_daily_ice %>% filter(waterYear == 2021) %>% filter(wy_doy > 183 & wy_doy <263)
daily_data_trimmed <- filter_by_year_and_doy(flow_temp_cond_daily_ice, c(183,288)) # April 1 - July 15
```

# Regression
```{r}
model1 <- glm(ice_or_no~Temperature_C+cond_uScm, data = flow_temp_cond_daily_ice, family = binomial)
summary(model1)

model1a <- glm(ice_or_no~Temperature_C+cond_uScm+Flow, data = flow_temp_cond_daily_ice, family = binomial)
summary(model1a)

model3 <- lm(ice_or_no~Temperature_C+cond_uScm, data = flow_temp_cond_daily_ice)
summary(model3)

model4 <- lm(ice_or_no~Temperature_C+cond_uScm+Flow, data = flow_temp_cond_daily_ice)
summary(model4)

model5 <- lm(ice_or_no~Temperature_C+cond_uScm+cumulative_dis, data = flow_temp_cond_daily_ice)
summary(model5)

plot(model1a)

model_window <- glm(ice_or_no~Temperature_C+cond_uScm+Flow, data = daily_data_trimmed, family = binomial)
summary(model_window)
```
# Binary Regressions following the Tutorial
```{r}
model_allDates <- glm(ice_presence~Temperature_C+cond_uScm+Flow, data = flow_temp_cond_daily_ice, family = binomial)
summary(model_allDates)

model1Var <- glm(ice_presence~Temperature_C, data = flow_temp_cond_daily_ice, family = binomial)
summary(model1Var)
exp(coef(model1Var)["Temperature_C"])
# the above exponential function of the coefficient for temperature means that one unit of increase with temperature increases the chances of there being no ice by a factor of 2.424
# the probability of there being no ice increases by 142% when temperature increases by 1

model1Var2021 <- glm(ic_presencee~Temperature_C, data = ALL_2021, family = binomial)
summary(model1Var2021)
exp(coef(model1Var2021)["Temperature_C"])
# in 2021, the probability of there being no ice increases by a factor of 117% when temperature increases by 1
```
# Temp Model Plot, Following Tutorial
```{r}
plot_model(model1Var,
  type = "pred",
  terms = "Temperature_C"
) +
  labs(y = "Prob (no ice)")
```
# Conductivity Model and Plot
```{r}
cond_model <- glm(ice_presence~cond_uScm, data = flow_temp_cond_daily_ice, family = binomial)
summary(cond_model)
exp(coef(cond_model)["cond_uScm"])
# -29% - probability of no ice decreases by 29% when conductivity rises by one unit (high conductivity = ice)
plot_model(cond_model,
  type = "pred",
  terms = "cond_uScm"
) +
  labs(y = "Prob (no ice)")
```
# Flow model and Plot
```{r}
flow_model <- glm(ice_presence~Flow, data = flow_temp_cond_daily_ice, family = binomial)
summary(flow_model)
exp(coef(flow_model)["Flow"])
# 21% - probability of no ice increases by 21% when Flow increases by one unit - high flow = no ice
plot_model(flow_model,
  type = "pred",
  terms = "Flow"
) +
  labs(y = "Prob (no ice)")
```
# Cumulative Flow Model and Plot
```{r}
cumul_flow_model <- glm(ice_presence~cumulative_dis, data = flow_temp_cond_daily_ice, family = binomial)
summary(cumul_flow_model)
exp(coef(cumul_flow_model)["cumulative_dis"])
# probability of no ice increases by < 1% when cumulative flow increases by one unit (cumulative flow units don't show relationship well)
plot_model(cumul_flow_model,
  type = "pred",
  terms = "cumulative_dis"
) +
  labs(y = "Prob (no ice)")
```
# All factors?
```{r}
all_model <- glm(ice_presence~Flow+Temperature_C+cond_uScm+waterYear, data = flow_temp_cond_daily_ice, family = binomial)
summary(all_model)

reduced_model <- glm(ice_presence~Temperature_C+cond_uScm+waterYear, data = flow_temp_cond_daily_ice, family = binomial)
# when just flow and water year - relationship between flow and ice is correct. 
# when flow and temp and water year or flow and cond and water year or flow and temp and cond or all three, relationship between flow and ice is wrong...
anova(reduced_model, all_model,
  test = "LRT"
)

# OR and 95% CI
round(exp(cbind(OR = coef(reduced_model), confint(reduced_model))), 3)
# only the ORs with significant p-values
round(exp(coef(reduced_model))[coef(summary(reduced_model))[, "Pr(>|z|)"] < 0.05], 3)
# probability of no ice decreases with increase in flow and conductivity, and increases with increase in temperature and water year -- high FLOW(?) and conductivity = ice, high temp and later year = no ice

plot_model(all_model,
  type = "pred",
  terms = c("Flow [all]","waterYear [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")

plot_model(reduced_model,
  type = "pred",
  terms = c("cond_uScm [all]","waterYear [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")

plot_model(reduced_model,
  type = "pred",
  terms = c("Temperature_C [all]","waterYear [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")
```
# Accuracy of temp, cond and water year
```{r}
# create a vector of predicted probabilities
preds1 <- predict(reduced_model,
  newdata = select(flow_temp_cond_daily_ice, -ice_presence), # remove real outcomes
  type = "response"
)

# if probability < threshold, ice IS on The Loch
preds_outcome1 <- ifelse(preds1 < 0.5,
  1,
  0
)

# transform predictions into factor and set labels
preds_outcome1 <- factor(preds_outcome1,
  levels = c(1, 0),
  labels = c("ice", "no ice")
)

# compare observed vs. predicted outcome
tab1 <- table(flow_temp_cond_daily_ice$ice_presence, preds_outcome1,
  dnn = c("observed", "predicted")
)

# print results
tab1

accuracy1 <- sum(diag(tab1)) / sum(tab1)
accuracy1

# sensitivity
sensitivity1 <- tab1[2, 2] / (tab1[2, 2] + tab1[2, 1])
sensitivity1

# specificity
specificity1 <- tab1[1, 1] / (tab1[1, 1] + tab1[1, 2])
specificity1
```
# Funciton for trimming dates each year
```{r}
# Define the function
filter_by_year_and_doy <- function(data, doy_range) {
  # Check if the doy_range is valid
  if (length(doy_range) != 2 || doy_range[1] >= doy_range[2]) {
    stop("Please provide a valid range in the form of c(start_doy, end_doy).")
  }

  # Create an empty data frame to store results
  result <- data.frame()

  # Iterate through each unique year
  unique_years <- unique(data$waterYear)
  
  for (year in unique_years) {
    # Filter the data for the current year and doy range
    filtered_data <- data %>%
      filter(waterYear == year, wy_doy >= doy_range[1], wy_doy <= doy_range[2])
    
    # Combine the filtered data into the result data frame
    result <- bind_rows(result, filtered_data)
  }
  
  return(result)
}



```
## Trimmed data models:
```{r}
trimmedDatModel <- glm(ice_presence~cumulative_dis+Temperature_C+cond_uScm+waterYear, data = daily_data_trimmed, family = binomial)
summary(trimmedDatModel)

reducedTrimModel <- glm(ice_presence~cumulative_dis+Temperature_C, data = daily_data_trimmed, family = binomial)
summary(reducedTrimModel)
```
## Trimmed Data Coef
```{r}
# OR and 95% CI
round(exp(cbind(OR = coef(reducedTrimModel), confint(reducedTrimModel))), 3)
# only the ORs with significant p-values
round(exp(coef(reducedTrimModel))[coef(summary(reducedTrimModel))[, "Pr(>|z|)"] < 0.05], 3)

plot_model(reducedTrimModel,
  type = "pred",
  terms = c("cumulative_dis [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")

plot_model(reducedTrimModel,
  type = "pred",
  terms = c("Temperature_C [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")
```


## Trimmed Data Accuracy
```{r}
# create a vector of predicted probabilities
predsTrim <- predict(reducedTrimModel,
  newdata = select(daily_data_trimmed, -ice_presence), # remove real outcomes
  type = "response"
)

# if probability < threshold, ice IS on The Loch
preds_outcomeTrim <- ifelse(predsTrim < 0.5,
  1,
  0
)

# transform predictions into factor and set labels
preds_outcomeTrim <- factor(preds_outcomeTrim,
  levels = c(1, 0),
  labels = c("ice", "no ice")
)

# compare observed vs. predicted outcome
tabTrim <- table(daily_data_trimmed$ice_presence, preds_outcomeTrim,
  dnn = c("observed", "predicted")
)

# print results
tabTrim

accuracyTrim <- sum(diag(tabTrim)) / sum(tabTrim)
accuracyTrim

# sensitivity
sensitivityTrim <- tabTrim[2, 2] / (tabTrim[2, 2] + tabTrim[2, 1])
sensitivityTrim

# specificity
specificityTrim <- tabTrim[1, 1] / (tabTrim[1, 1] + tabTrim[1, 2])
specificityTrim
```
# AUC (trimmed daily)
```{r}
aucTrim <- roc(ice_presence ~ fitted(reducedTrimModel),
  data = daily_data_trimmed)
ggroc(aucTrim, legacy.axes = TRUE) +
  labs(title = paste0("AUC = ", round(aucTrim$auc, 2)))
```
# Summary
```{r}
tbl_regression(reducedTrimModel, exponentiate = TRUE)
```



## Trimmed Data with Predictions
```{r}
predValuesAll <- predict(reducedTrimModel, type = "response")
dfPred <- data.frame(observed = daily_data_trimmed$ice_presence, predicted = predValuesAll)
ggplot(dfPred, aes(y = predicted, x = observed))+
  geom_point()
```
## Using Predict Function to predict values on entire dataset
```{r}
testAllPredict <- flow_temp_cond_daily_ice
# Generate predictions for the new dataset
testAllPredict$predicted_prob_ice <- predict(reducedTrimModel, newdata = testAllPredict, type = "response")
# Apply threshold for binary predicitons
testAllPredict$predicted_ice_presence <- ifelse(testAllPredict$predicted_prob_ice >= 0.5, 0,1)

View(testAllPredict)

```


# Testing interaction
```{r}
inter_model <- glm(ice_presence~Temperature_C*cond_uScm*Flow*waterYear, data = flow_temp_cond_daily_ice, family = binomial)
summary(inter_model)
anova(reduced_model, inter_model,
  test = "LRT"
)
round(exp(coef(inter_model))[coef(summary(inter_model))[, "Pr(>|z|)"] < 0.05], 3)

inter_model2 <- glm(ice_presence~Temperature_C+cond_uScm*Flow*waterYear, data = flow_temp_cond_daily_ice, family = binomial)
summary(inter_model2)

round(exp(coef(inter_model2))[coef(summary(inter_model2))[, "Pr(>|z|)"] < 0.05], 3)

anova(inter_model2, inter_model,
  test = "LRT"
)

inter_model3 <- glm(ice_presence~Temperature_C+cond_uScm*Flow+waterYear, data = flow_temp_cond_daily_ice, family = binomial)
summary(inter_model3)

round(exp(coef(inter_model3))[coef(summary(inter_model3))[, "Pr(>|z|)"] < 0.05], 3)

anova(inter_model3, inter_model2,
  test = "LRT"
)

plot_model(inter_model3,
  type = "pred",
  terms = c("Flow [all]","waterYear [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")
# what the heck is this

inter_model4 <- glm(ice_presence~Temperature_C*cond_uScm*waterYear, data = flow_temp_cond_daily_ice, family = binomial)
summary(inter_model4)

round(exp(coef(inter_model4))[coef(summary(inter_model4))[, "Pr(>|z|)"] < 0.05], 3)

anova(inter_model4, inter_model3,
  test = "LRT"
)

inter_model5 <- glm(ice_presence~Temperature_C+cond_uScm*waterYear, data = flow_temp_cond_daily_ice, family = binomial)
summary(inter_model5)

round(exp(coef(inter_model5))[coef(summary(inter_model5))[, "Pr(>|z|)"] < 0.05], 3)

anova(inter_model5, inter_model4,
  test = "LRT")

plot_model(inter_model5,
  type = "pred",
  terms = c("Temperature_C [all]","waterYear [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")

plot_model(inter_model5,
  type = "pred",
  terms = c("cond_uScm","waterYear [all]"),
  ci.lvl = NA # remove confidence bands
) +
  labs(y = "Prob (no ice)")

```
# Model selection
```{r}
# save initial model
all_variables_m <- glm(ice_presence ~ (Flow+cumulative_dis+Temperature_C+cond_uScm+waterYear)^2,
  data = flow_temp_cond_daily_ice,
  family = "binomial"
)
# select best model according to AIC using mixed selection
AIC_chosen_model <- step(all_variables_m,
  direction = "both", # both = mixed selection
  trace = FALSE # do not display intermediate steps
)
# display results of final model
summary(AIC_chosen_model)
```
# Testing Predictions
```{r}
# create a vector of predicted probabilities
preds <- predict(AIC_chosen_model,
  newdata = select(flow_temp_cond_daily_ice, -ice_presence), # remove real outcomes
  type = "response"
)

# if probability < threshold, ice IS on The Loch
preds_outcome <- ifelse(preds < 0.5,
  1,
  0
)

# transform predictions into factor and set labels
preds_outcome <- factor(preds_outcome,
  levels = c(1, 0),
  labels = c("ice", "no ice")
)

# compare observed vs. predicted outcome
tab <- table(flow_temp_cond_daily_ice$ice_presence, preds_outcome,
  dnn = c("observed", "predicted")
)

# print results
tab

accuracy <- sum(diag(tab)) / sum(tab)
accuracy

# sensitivity
sensitivity <- tab[2, 2] / (tab[2, 2] + tab[2, 1])
sensitivity

# specificity
specificity <- tab[1, 1] / (tab[1, 1] + tab[1, 2])
specificity
```
# AUC and ROC
```{r}
# save roc object
res <- roc(ice_presence ~ fitted(AIC_chosen_model),
  data = flow_temp_cond_daily_ice
)

# plot ROC curve
ggroc(res, legacy.axes = TRUE)
```


```{r}
pred_values <- predict(model_window, type = "response")
df <- data.frame(observed = daily_data_trimmed$ice_or_no, predicted = pred_values)
ggplot(df, aes(y = predicted, x = observed))+
  geom_point()
```

